---
title: "Data Collection"
author: "Yueyi Xu"
date: "2023-12-04"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(readxl)
library(openxlsx)
```

### Clean the dataset for the top 25 amusement/theme parks Worldwide in 2019 to 2022
```{r}
top25_tp_worldwide =
  read_excel("statistic_id194247_most-visited-amusement-and-theme-parks-worldwide-2019-2022.xlsx", sheet = "Data")

top25_tp_worldwide_clean_df =
  top25_tp_worldwide |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  separate(
    col = ("Most visited amusement and theme parks worldwide 2019-2022"), into = c("Park_Name", "Country_State"), sep = ","
  ) |>
  mutate(
    Type = "Amusement/Theme Park",
    Region = "Worldwide",
    Attendance = Attendance * 10e6
  ) |>
  select(
    Park_Name, Country_State, Year, Attendance, Type, Region
  )
```


### Clean the dataset for the top 20 amusement/theme parks North America in 2019 to 2022
```{r}
top20_tp_na =
  read_excel("statistic_id194269_most-visited-amusement-and-theme-parks-in-north-america-2022.xlsx", sheet = "Data")

top20_tp_na_clean_df =
  top20_tp_na |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  separate(
    col = ("Most visited amusement and theme parks in North America 2022"), into = c("Park_Name", "Country_State"), sep = ","
  ) |>
  mutate(
    Type = "Amusement/Theme Park",
    Region = "North America",
    Attendance = Attendance * 10e6
  ) |>
  select(
    Park_Name, Country_State, Year, Attendance, Type, Region
  )
```


### Clean the dataset for the top 10 amusement/theme parks Latin America in 2019 to 2022
*8 total
```{r}
top10_tp_la =
  read_excel("statistic_id194312_most-visited-latin-american-amusement-and-theme-parks-2019-2022.xlsx", sheet = "Data")

top10_tp_la_clean_df =
  top10_tp_la |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  separate(
    col = ("Most visited Latin American amusement and theme parks 2019-2022"), into = c("Park_Name", "Country_State"), sep = ","
  ) |>
  mutate(
    Type = "Amusement/Theme Park",
    Region = "Latin America",
    Attendance = Attendance * 10e6
  ) |>
  select(
    Park_Name, Country_State, Year, Attendance, Type, Region
  )
```


### Clean the dataset of the top 20 Amusement/Theme Parks Asia Pacific in 2019 to 2021
*Fix (city+country)? and sinapore without country name
```{r}
top20_tp_apac =
  read_excel("statistic_id194300_most-visited-amusement-and-theme-parks-apac-2019-2021.xlsx", sheet = "Data")

top20_tp_apac_clean_df =
  top20_tp_apac |>
  pivot_longer(
    ("2019"):("2021"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  separate(
    col = ("Most visited amusement and theme parks APAC 2019-2021"), into = c("Park_Name", "Country_State"), sep = ","
  ) |>
  mutate(
    Type = "Amusement/Theme Park",
    Region = "Asia Pacific",
    Attendance = Attendance * 10e6
  ) |>
  select(
    Park_Name, Country_State, Year, Attendance, Type, Region
  )
```


### Clean the dataset for the top 20 water parks North America in 2019 to 2022
```{r}
top20_wp_na =
  read_excel("statistic_id194352_total-attendance-at-waterparks-in-the-us-2019-2022.xlsx", sheet = "Data")

top20_wp_na_clean_df =
  top20_wp_na |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  separate(
    col = ("Total attendance at waterparks in the U.S. 2019-2022"), into = c("Park_Name", "Country_State"), sep = ","
  ) |>
  mutate(
    Type = "Water Park",
    Region = "North America",
    Attendance = Attendance * 10e6
  ) |>
  select(
    Park_Name, Country_State, Year, Attendance, Type, Region
  )
```


### Clean the dataset for the top 10 water parks Latin America in 2020
```{r}
top10_wp_la =
  read_excel("statistic_id1024329_most-visited-latin-american-water-parks-2020.xlsx", sheet = "Data")

top10_wp_la_clean_df =
  top10_wp_la |>
  pivot_longer(
    ("2020"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  separate(
    col = ("Most visited Latin American water parks 2020"), into = c("Park_Name", "Country_State"), sep = ","
  ) |>
  mutate(
    Type = "Water Park",
    Region = "Latin America",
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, Country_State, Year, Attendance, Type, Region
  )
```


### Clean the dataset for the top 20 water parks Asia Pacific in 2019 to 2021
```{r}
top20_wp_apac =
  read_excel("statistic_id1346887_most-visited-water-parks-apac-2019-2021.xlsx", sheet = "Data")

top20_wp_apac_clean_df =
  top20_wp_apac |>
  pivot_longer(
    ("2019"):("2021"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  separate(
    col = ("Most visited water parks APAC 2019-2021"), into = c("Park_Name", "Country_State"), sep = ","
  ) |>
  mutate(
    Type = "Water Park",
    Region = "Asia Pacific",
    Attendance = Attendance * 10e6
  ) |>
  select(
    Park_Name, Country_State, Year, Attendance, Type, Region
  )
```


### Clean the dataset for the top 20 museums Worldwide in 2019 to 2022
*19 total
```{r}
top20_mu_worldwide =
  read_excel("statistic_id901072_leading-museums-by-highest-attendance-worldwide-2019-2022.xlsx", sheet = "Data")

top20_mu_worldwide_clean_df =
  top20_mu_worldwide |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  separate(
    col = ("Leading museums by highest attendance worldwide 2019-2022"), into = c("Park_Name", "Country_State"), sep = ","
  ) |>
  mutate(
    Type = "Museum",
    Region = "Worldwide",
    Attendance = Attendance * 10e6
  ) |>
  select(
    Park_Name, Country_State, Year, Attendance, Type, Region
  )
```


### Clean the dataset for the top 20 museums Europe Middle East Africa in 2019 to 2022
```{r}
top20_mu_emea =
  read_excel("statistic_id747942_most-visited-museums-in-europe-2019-2022.xlsx", sheet = "Data")

top20_mu_emea_clean_df =
  top20_mu_emea |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  separate(
    col = ("Most visited museums in Europe 2019-2022"), into = c("Park_Name", "Country_State"), sep = ","
  ) |>
  mutate(
    Type = "Museum",
    Region = "Europe",
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, Country_State, Year, Attendance, Type, Region
  )
```


### Import the amusement/theme parks data of EMEA
```{r}
data_text_tp_emea = "1,DISNEYLAND PARK AT DISNEYLAND PARIS, MARNE-LA-VALLEE, FRANCE,184%,9930,3500,2620,9745
2,EUROPA-PARK, RUST, GERMANY,80%,5400,3000,2500,5750
3,DE EFTELING THEME PARK RESORT, KAATSHEUVEL, NETHERLANDS,65%,5430,3300,2900,5260
4,WALT DISNEY STUDIOS PARK AT DISNEYLAND PARIS, MARNE-LA-VALLEE, FRANCE,183%,5340,1884,1410,5245
5,TIVOLI GARDENS, COPENHAGEN, DENMARK,61%,3864,2400,1628,4581
6,PORTAVENTURA WORLD, SALOU, SPAIN,56%,3750,2400,700,3750
7,LISEBERG, GOTHENBURG, SWEDEN,73%,2500,1447,0,2950
8,GARDALAND, CASTELNUOVO DEL GARDA, ITALY,34%,2950,2200,1350,2920
9,LEGOLAND WINDSOR, WINDSOR, U.K.,60%,2400,1500,450,2430
10,PARC ASTERIX, PLAILLY, FRANCE,102%,2632,1300,1163,2326
11,PUY DU FOU, LES EPESSES, FRANCE,45%,2342,1616,923,2308
12,PARQUE WARNER, MADRID, SPAIN,43%,1860,1300,450,2232
13,ALTON TOWERS, STAFFORDSHIRE, U.K.,28%,2300,1800,670,2130
14,PHANTASIALAND, BRÜHL, GERMANY,78%,2100,1180,1000,2050
15,LEGOLAND BILLUND, BILLUND, DENMARK,164%,2242,850,700,1950
16,THORPE PARK, CHERTSEY, U.K.,-6%,1600,1700,600,1900
17,FUTUROSCOPE, JAUNAY-CLAN, FRANCE,75%,1920,1100,900,1900
18,LEGOLAND DEUTSCHLAND, GÜNZBURG, GERMANY,89%,1700,900,750,1700
19,HEIDE PARK, SOLTAU, GERMANY,23%,1600,1300,950,1700
20,CHESSINGTON WORLD OF ADVENTURES, CHESSINGTON, U.K.,3%,1500,1450,510,1690"

# Read the corrected data into a data frame
top20_tp_emea = read.table(text = data_text_tp_emea, sep = ",", header = FALSE, quote = "", na.strings = "-")
colnames(top20_tp_emea) = c("id","Park_Name", "City", "Country", "%Change", "2022", "2021", "2020", "2019")
```


### Clean the dataset for the top 20 amusement/theme parks Europe Middle East Africa in 2019 to 2022
```{r}
top20_tp_emea_clean_df =
  top20_tp_emea |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Amusement/Theme Park",
    Region = "EMEA",
    Country_State = Country,
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, Country_State, Year, Attendance, Type, Region
  )
```


### Import the water parks data of EMEA
```{r}
data_text_wp_emea = "1, THERME ERDING, ERDING, GERMANY, 143%, 1700, 700, 750, 1850
2, AQUAVENTURE WATER PARK, DUBAI, U.A.E., 88%, 1500, 800, 600, 1322
3, AQUAPALACE, PRAGUE, CZECH REPUBLIC, 107%, 1200, 580, 537, 1300
4, TROPICAL ISLANDS, KRAUSNICK, GERMANY, 130%, 1150, 500, 493, 1233
5, SIAM PARK, SANTA CRUZ DE TENERIFE, SPAIN, 125%, 1150, 511, 97, 1200
6, AQUALAND MORAVIA, PASOHLÁVKY, CZECH REPUBLIC, 88%, 727, 386, 368, 806
7, TIKI POOL, DUINRELL, NETHERLANDS, 88%, 750, 400, 390, 800
8, NETTEBAD, OSNABRÜCK, GERMANY, 87%, 650, 347, 311, 756
9, WILD WADI, DUBAI, U.A.E., 60%, 750, 470, 350, 740
10, LALANDIA, BILLUND, DENMARK, 67%, 750, 450, 240, 682"
top10_wp_emea = read.table(text = data_text_wp_emea, sep = ",", header = FALSE, quote = "")
colnames(top10_wp_emea) = c("id","Park_Name", "City", "Country", "%Change", "2022", "2021", "2020", "2019")
```


### Clean the dataset for the top 10 water parks Europe Middle East Africa in 2019 to 2022
```{r}
top10_wp_emea_clean_df =
  top10_wp_emea |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Water Park",
    Region = "EMEA",
    Country_State = Country,
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, Country_State, Year, Attendance, Type, Region
  )
```



### Combine all dataset
```{r}
combine_df = 
  bind_rows(top25_tp_worldwide_clean_df, top20_tp_na_clean_df, top10_tp_la_clean_df, top20_tp_apac_clean_df, top20_wp_na_clean_df, top10_wp_la_clean_df, top20_wp_apac_clean_df, top20_mu_worldwide_clean_df, top20_mu_emea_clean_df)
```


### Export the final dataset
```{r}
write.csv(combine_df, "final data.csv", row.names = FALSE)
```





