---
title: "Data Collection_Updated"
author: "Yueyi Xu"
date: "2023-12-05"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    code_folding: hide
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(readxl)
library(openxlsx)
```

# Data Collection and Data Source
Our data was collected from the “Global Attractions Attendance Report” which was published by AECOM and TEA. We included the attendance data of different Theme parks and Museums worldwide from 2019 to 2022. The article can be found here.
https://aecom.com/wp-content/uploads/documents/reports/AECOM-Theme-Index-2022.pdf

### Data Import and Cleaning
Each individual dataset catgegorized by types of parks and regions of the parks are listed in the article. By collecting each of the dataset, we created separate dataframes and combine to one ultimate dataset with all of the data focusing on "Park_Name", "City", "Country", "Year", "Attendance", "Type", and "Region".

"Park_Name": Names of the park
"City": City of where the park is located
"Country": Country of where the park is located
"Year": Year of attendance
"Attendance": Number of people visited the park
"Type": Amusement/Theme Parks, Water Parks, and Museums
"Region": Worldwide, North America, Latin America, Asia Pacific, Europe Middle East Africa


### Import the amusement/theme parks data of Worldwide 2019-2022
```{r, echo=FALSE}
data_text_tp_ww = "1, MAGIC KINGDOM THEME PARK AT WALT DISNEY WORLD RESORT, LAKE BUENA VISTA, U.S., 35%, 17133, 12691, 6941, 20963
2, DISNEYLAND PARK AT DISNEYLAND RESORT, ANAHEIM, U.S., 97%, 16881, 8573, 3674, 18666
3, TOKYO DISNEYLAND AT TOKYO DISNEY RESORT, TOKYO, JAPAN, 90%, 12000, 6300, 4160, 17910
4, TOKYO DISNEYSEA AT TOKYO DISNEY RESORT, TOKYO, JAPAN, 74%, 10100, 5800, 3400, 14650
5, UNIVERSAL STUDIOS JAPAN, OSAKA, JAPAN, 125%, 12350, 5500, 4901, 14500
6, DISNEY'S ANIMAL KINGDOM AT WALT DISNEY WORLD, LAKE BUENA VISTA, U.S., 25%, 9027, 7194, 4166, 13888
7, EPCOT AT WALT DISNEY WORLD, LAKE BUENA VISTA, U.S., 29%, 10000, 7752, 4044, 12444
8, CHIMELONG OCEAN KINGDOM, HENGQIN, CHINA, -41%, 4400, 7452, 4797, 11736
9, DISNEY'S HOLLYWOOD STUDIOS AT WALT DISNEY WORLD, LAKE BUENA VISTA, U.S., 27%, 10900, 8589, 3675, 11483
10, SHANGHAI DISNEYLAND AT SHANGHAI DISNEY RESORT, SHANGHAI, CHINA, -38%, 5300, 8480, 5500, 11210
11, UNIVERSAL STUDIOS FLORIDA AT UNIVERSAL ORLANDO, FL, U.S., 20%, 10750, 8987, 4096, 10922
12, UNIVERSAL'S ISLANDS OF ADVENTURE AT UNIVERSAL ORLANDO, ORLANDO, U.S., 21%, 11025, 9077, 4005, 10375
13, DISNEYCALIFORNIA ADVENTURE PARK AT DISNEYLAND RESORT, ANAHEIM, U.S., 81%, 9000, 4977, 1919, 9861
14, DISNEYLAND PARK AT DISNEYLAND PARIS, MARNE-LA-VALLEE, FRANCE, 184%, 9930, 3500, 2620, 9745
15, UNIVERSAL STUDIOS HOLLYWOOD, UNIVERSAL CITY, U.S., 53%, 8400, 5505, 1299, 9147
16, EVERLAND, GYEONGGI-DO, SOUTH KOREA, 56%, 5770, 3710, 2760, 6606
17, LOTTE WORLD, SEOUL, SOUTH KOREA, 84%, 4520, 2460, 1560, 5953
18, NAGASHIMA SPA LAND, KUWANA, JAPAN, 17%, 4200, 3600, 2400, 5950
19, EUROPA-PARK, RUST, GERMANY, 100%, 6000, 3000, 2500, 5750
20, OCEAN PARK, HONG KONG, SAR, 0%, 1400, 1400, 2200, 5700
21, HONG KONG DISNEYLAND, HONG KONG, SAR, 21%, 3400, 2800, 1700, 5695
22, DE EFTELING EFTELING THEME PARK RESORT, KAATSHEUVEL, NETHERLANDS, 65%, 5430, 3300, 2900, 5260
23, WALT DISNEY STUDIOS PARK AT DISNEYLAND PARIS, MARNE-LA-VALLEE, FRANCE, 183%, 5340, 1884, 1410, 5245
24, OCT HAPPY VALLEY, BEIJING, CHINA, -24%, 3740, 4930, 3950, 5160
25, CHIMELONG PARADISE, GUANGZHOU, CHINA, -41%, 2300, 3890, 2681, 4905"

top25_tp_ww = read.table(text = data_text_tp_ww, sep = ",", header = FALSE, quote = "", na.strings = "-")
colnames(top25_tp_ww) = c("id","Park_Name", "City", "Country", "%Change", "2022", "2021", "2020", "2019")
```


### Clean the dataset for the top 25 amusement/theme parks Worldwide in 2019 to 2022
```{r, echo=TRUE}
top25_tp_ww_clean_df =
  top25_tp_ww |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Amusement/Theme Park",
    Region = "Worldwide",
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```


### Import the amusement/theme parks data of North America 2019-2022
```{r, echo=FALSE}
data_text_tp_na = "1, MAGIC KINGDOM THEME PARK AT WALT DISNEY WORLD, LAKE BUENA VISTA, FL, U.S., 35%, 17133, 12691, 6941, 20963
2, DISNEYLAND PARK AT DISNEYLAND RESORT, ANAHEIM, CA, U.S., 97%, 16881, 8573, 3674, 18666
3, DISNEY'S ANIMAL KINGDOM AT WALT DISNEY WORLD, LAKE BUENA VISTA, FL, U.S., 25%, 9027, 7194, 4166, 13888
4, EPCOT AT WALT DISNEY WORLD, LAKE BUENA VISTA, FL, U.S., 29%, 10000, 7752, 4044, 12444
5, DISNEY'S HOLLYWOOD STUDIOS AT WALT DISNEY WORLD, LAKE BUENA VISTA, FL, U.S., 27%, 10900, 8589, 3675, 11483
6, UNIVERSAL STUDIOS FLORIDA, ORLANDO, FL, U.S., 20%, 10750, 8987, 4096, 10922
7, UNIVERSAL'S ISLANDS OF ADVENTURE, ORLANDO, FL, U.S., 21%, 11025, 9077, 4005, 10375
8, DISNEY CALIFORNIA ADVENTURE PARK AT DISNEYLAND RESORT, ANAHEIM, CA, U.S., 81%, 9000, 4977, 1919, 9861
9, UNIVERSAL STUDIOS HOLLYWOOD, UNIVERSAL CITY, CA, U.S., 53%, 8400, 5505, 1299, 9147
10, SEAWORLD ORLANDO, ORLANDO, FL, U.S., 46%, 4454, 3051, 1598, 4640
11, KNOTT'S BERRY FARM, BUENA PARK, CA, U.S., 6%, 3899, 3681, 811, 4238
12, BUSCH GARDENS TAMPA BAY, TAMPA, FL, U.S., 26%, 4051, 3210, 1288, 4180
13, CANADA'S WONDERLAND, MAPLE, ONTARIO, CANADA, 524%, 3768, 587, 0, 3950
14, SEAWORLD SAN DIEGO, SAN DIEGO, CA, U.S., 25%, 3507, 2800, 1139, 3731
15, CEDAR POINT, SANDUSKY, OH, U.S., 4%, 3444, 3327, 1020, 3610
16, SIX FLAGS MAGIC MOUNTAIN, VALENCIA, CA, U.S., -2%, 2993, 3047, 686, 3521
17, KINGS ISLAND, KINGS ISLAND, OH, U.S., 5%, 3340, 3181, 1626, 3485
18, SIX FLAGS GREAT ADVENTURE, JACKSON, NJ, U.S., -26%, 2153, 2913, 598, 3451
19, HERSHEYPARK, HERSHEY, PA, U.S., 6%, 3193, 3012, 1717, 3384
20, SIX FLAGS GREAT AMERICA, GURNEE, IL, U.S., -5%, 2535, 2675, 0, 3169"

top20_tp_na = read.table(text = data_text_tp_na, sep = ",", header = FALSE, quote = "")
colnames(top20_tp_na) = c("id","Park_Name", "City", "State", "Country", "%Change", "2022", "2021", "2020", "2019")
```


### Clean the dataset for the top 20 amusement/theme parks North America in 2019 to 2022
```{r, echo=TRUE}
top20_tp_na_clean_df =
  top20_tp_na |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Amusement/Theme Park",
    Region = "North America",
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```


### Import the amusement/theme parks data of Latin America 2019-2022
```{r, echo=FALSE}
data_text_tp_la = "1,SIX FLAGS MEXICO, MEXICO CITY, MEXICO, 54%, 1727, 1125, 701, 2803
2,BETO CARRERO WORLD, SANTA CATARINA, BRAZIL, 23%, 2327, 1895, 1252, 2241
3,PARQUE XCARET, CANCUN, MEXICO, 27%, 1399, 1104, 736, 1960
4,MUNDO PETAPA, GUATEMALA CITY, GUATEMALA, 99%, 800, 402, 268, 1330
5,LA FERIA DE CHAPULTEPEC, MEXICO CITY, MEXICO, 0%, 0, 0, 0, 1306
6,PLAZA DE SESAMO, MONTERREY, MEXICO, 0%, 0, 0, 0, 1173
7,PARQUE MUNDO AVENTURA, BOGOTÁ, COLOMBIA, 19%, 752, 631, 344, 1151
8,FANTASIALANDIA, SANTIAGO, CHILE, 38%, 850, 615, 430, 1100
9,THEME PARQUE NACIONAL DEL CAFÉ, QUINDIO, COLOMBIA, 30%, 626, 482, 321, 1043
10,PARQUE DE LA COSTA, BUENOS AIRES, ARGENTINA, 30%, 514, 395, 263, 968"

top10_tp_la = read.table(text = data_text_tp_la, sep = ",", header = FALSE, quote = "")
colnames(top10_tp_la) = c("id","Park_Name", "City", "Country", "%Change", "2022", "2021", "2020", "2019")
```


### Clean the dataset for the top 10 amusement/theme parks Latin America in 2019 to 2022
```{r, echo=TRUE}
top10_tp_la_clean_df =
  top10_tp_la |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Amusement/Theme Park",
    Region = "Latin America",
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```


### Import the amusement/theme parks data of Asia Pacific 2019-2022
```{r, echo=FALSE}
data_text_tp_apac = "1, TOKYO DISNEYLAND AT TOKYO DISNEY RESORT, TOKYO, JAPAN, 90%, 12000, 6300, 4160, 17910
2, TOKYO DISNEYSEA AT TOKYO DISNEY RESORT, TOKYO, JAPAN, 74%, 10100, 5800, 3400, 14650
3, UNIVERSAL STUDIOS JAPAN, OSAKA, JAPAN, 125%, 12350, 5500, 4901, 14500
4, CHIMELONG OCEAN KINGDOM, HENGQIN, CHINA, -41%, 4400, 7452, 4797, 11736
5, SHANGHAI DISNEYLAND AT SHANGHAI DISNEY RESORT, SHANGHAI, CHINA, -38%, 5300, 8480, 5500, 11210
6, EVERLAND, SEOUL, SOUTH KOREA, 56%, 5770, 3710, 2760, 6606
7, LOTTE WORLD, SEOUL, SOUTH KOREA, 84%, 4520, 2460, 1560, 5953
8, NAGASHIMA SPA LAND, KUWANA, JAPAN, 17%, 4200, 3600, 2400, 5950
9, OCEAN PARK, HONG KONG, SAR, 0%, 1400, 1400, 2200, 5700
10, HONG KONG DISNEYLAND, HONG KONG, SAR, 21%, 3400, 2800, 1700, 5695
11, OCT HAPPY VALLEY, BEIJING, CHINA, -24%, 3740, 4930, 3950, 5160
12, CHIMELONG PARADISE, GUANGZHOU, CHINA, -41%, 2300, 3890, 2681, 4905
13, UNIVERSAL STUDIOS SINGAPORE, Sentosa, SINGAPORE, 75%, 2100, 1200, 1098, 4500
14, CHINA DINOSAUR PARK, CHANGZHOU, CHINA, -12%, 2200, 2500, 2375, 4434
15, OCT WINDOW OF THE WORLD, SHENZHEN, CHINA, -54%, 1100, 2400, 1890, 3990
16, OCT HAPPY VALLEY, SHENZHEN, CHINA, -47%, 1750, 3310, 3120, 3980
17, ZHENGZHOU FANTAWILD ADVENTURE, ZHENGZHOU, CHINA, -42%, 1590, 2752, 3421, 3840
18, OCT HAPPY VALLEY, CHENGDU, CHINA, -47%, 1570, 2940, 2620, 3580
19, NINGBO FANTAWILD ORIENTAL HERITAGE, NINGBO, CHINA, -47%, 1230, 2308, 2257, 3575
20, OCT HAPPY VALLEY, SHANGHAI, CHINA, -62%, 1240, 3300, 2730, 3390"

top20_tp_apac = read.table(text = data_text_tp_apac, sep = ",", header = FALSE, quote = "", na.strings = "-")
colnames(top20_tp_apac) = c("id","Park_Name", "City", "Country", "%Change", "2022", "2021", "2020", "2019")
```


### Clean the dataset of the top 20 Amusement/Theme Parks Asia Pacific in 2019 to 2022
```{r, echo=TRUE}
top20_tp_apac_clean_df =
  top20_tp_apac |>
  pivot_longer(
    ("2019"):("2021"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Amusement/Theme Park",
    Region = "Asia Pacific",
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```


### Import the amusement/theme parks data of EMEA
```{r, echo=FALSE}
data_text_tp_emea = "1,DISNEYLAND PARK AT DISNEYLAND PARIS, MARNE-LA-VALLEE, FRANCE,184%,9930,3500,2620,9745
2,EUROPA-PARK, RUST, GERMANY,80%,5400,3000,2500,5750
3,DE EFTELING THEME PARK RESORT, KAATSHEUVEL, NETHERLANDS,65%,5430,3300,2900,5260
4,WALT DISNEY STUDIOS PARK AT DISNEYLAND PARIS, MARNE-LA-VALLEE, FRANCE,183%,5340,1884,1410,5245
5,TIVOLI GARDENS, COPENHAGEN, DENMARK,61%,3864,2400,1628,4581
6,PORTAVENTURA WORLD, SALOU, SPAIN,56%,3750,2400,700,3750
7,LISEBERG, GOTHENBURG, SWEDEN,73%,2500,1447,0,2950
8,GARDALAND, CASTELNUOVO DEL GARDA, ITALY,34%,2950,2200,1350,2920
9,LEGOLAND WINDSOR, WINDSOR, U.K.,60%,2400,1500,450,2430
10,PARC ASTERIX, PLAILLY, FRANCE,102%,2632,1300,1163,2326
11,PUY DU FOU, LES EPESSES, FRANCE,45%,2342,1616,923,2308
12,PARQUE WARNER, MADRID, SPAIN,43%,1860,1300,450,2232
13,ALTON TOWERS, STAFFORDSHIRE, U.K.,28%,2300,1800,670,2130
14,PHANTASIALAND, BRÜHL, GERMANY,78%,2100,1180,1000,2050
15,LEGOLAND BILLUND, BILLUND, DENMARK,164%,2242,850,700,1950
16,THORPE PARK, CHERTSEY, U.K.,-6%,1600,1700,600,1900
17,FUTUROSCOPE, JAUNAY-CLAN, FRANCE,75%,1920,1100,900,1900
18,LEGOLAND DEUTSCHLAND, GÜNZBURG, GERMANY,89%,1700,900,750,1700
19,HEIDE PARK, SOLTAU, GERMANY,23%,1600,1300,950,1700
20,CHESSINGTON WORLD OF ADVENTURES, CHESSINGTON, U.K.,3%,1500,1450,510,1690"

top20_tp_emea = read.table(text = data_text_tp_emea, sep = ",", header = FALSE, quote = "", na.strings = "-")
colnames(top20_tp_emea) = c("id","Park_Name", "City", "Country", "%Change", "2022", "2021", "2020", "2019")
```


### Clean the dataset for the top 20 amusement/theme parks Europe Middle East Africa in 2019 to 2022
```{r, echo=TRUE}
top20_tp_emea_clean_df =
  top20_tp_emea |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Amusement/Theme Park",
    Region = "EMEA",
    Country_State = Country,
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```


### Import water parks data of North America 2019-2022
```{r, echo=FALSE}
data_text_wp_na = "1,DISNEY'S TYPHOON LAGOON WATER PARK AT WALT DISNEY WORLD RESORT, ORLANDO, FL, U.S., 0%, 1915, 0, 0, 2248
2,DISNEY'S BLIZZARD BEACH WATER PARK AT WALT DISNEY WORLD RESORT, ORLANDO, FL, U.S., -92%, 101, 1201, 316, 1983
3,UNIVERSAL'S VOLCANO BAY, ORLANDO, FL, U.S., 9%, 1850, 1691, 551, 1811
4,AQUATICA, ORLANDO, FL, U.S., 35%, 1548, 1147, 528, 1533
5,SCHLITTERBAHN, NEW BRAUNFELS, TX, U.S., 7%, 946, 888, 451, 996
6,WATER COUNTRY USA, WILLIAMSBURG, VA, U.S., 1%, 707, 700, 0, 736
7,ADVENTURE ISLAND, TAMPA, FL, U.S., 7%, 636, 593, 125, 656
8,AQUATICA SAN ANTONIO, SAN ANTONIO, TX, U.S., 8%, 631, 586, 263, 651
9,SCHLITTERBAHN, GALVESTON, TX, U.S., 34%, 539, 402, 337, 567
10,SPLISH SPLASH, CALVERTON, NY, U.S., 0, 0, 0, 0, 542
11,SIX FLAGS HURRICANE HARBOR, ARLINGTON, TX, U.S., 3%, 388, 377, 235, 538
12,SIX FLAGS WHITE WATER ATLANTA, MARIETTA, GA, U.S., -5%, 350, 368, 238, 526
13,SIX FLAGS HURRICANE HARBOR, JACKSON, NJ, U.S., -5%, 426, 448, 85, 482
14,TYPHOON TEXAS, KATY, TX, U.S., 7%, 364, 340, 203, 478
15,SIX FLAGS HURRICANE HARBOR SPLASHTOWN, SPRING, TX, U.S., 3%, 343, 333, 0, 475
16,ZOOMBEZI BAY, POWELL, OH, U.S., 7%, 399, 373, 176, 466
17,SIX FLAGS HURRICANE HARBOR PHOENIX, GLENDALE, AZ, U.S., 3%, 318, 309, 0, 441
18,CAMELBEACH, TANNERSVILLE, PA, U.S., 66%, 373, 225, 151, 439
19,DOLLYWOOD'S SPLASH COUNTRY, PIGEON FORGE, TN, U.S., 31%, 368, 281, 188, 433
20,CEDAR POINT SHORES, SANDUSKY, OH, U.S., 1%, 405, 402, 0, 424"

top20_wp_na = read.table(text = data_text_wp_na, sep = ",", header = FALSE, quote = "")
colnames(top20_wp_na) = c("id","Park_Name", "City", "State", "Country", "%Change", "2022", "2021", "2020", "2019")
```


### Clean the dataset for the top 20 water parks North America in 2019 to 2022
```{r, echo=TRUE}
top20_wp_na_clean_df =
  top20_wp_na |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Water Park",
    Region = "North America",
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```


### Import water parks data of Latin America 2019-2022
```{r, echo=FALSE}
data_text_wp_la = "1, THERMAS DOS LARANJAIS, OLIMPIA, BRAZIL, 47%, 1729, 1179, 922, 1845
2, BAHAMAS AQUAVENTURE WATER PARK, PARADISE ISLAND (NASSAU NEW PROVIDENCE), BAHAMAS, 38%, 1751, 1269, 494, 1813
3, HOT PARK RIO QUENTE, RIO QUENTE, BRAZIL, 53%, 1430, 933, 734, 1469
4, PARQUE ACUATICO XOCOMIL, SAN MARTÍN ZAPOTITLÁN, GUATEMALA, 33%, 833, 626, 121, 894
5, BEACH PARK, AQUIRAZ, BRAZIL, 56%, 782, 500, 312, 890
6, PISCILAGO, GIRARDOT (BOGOTÁ), COLOMBIA, 34%, 816, 609, 307, 876
7, MAGIC CITY WATER PARK, SUZANO, BRAZIL, 32%, 616, 467, 355, 681
8, THERMAS WATER PARK, SÃO PEDRO, BRAZIL, 67%, 753, 452, 315, 641
9, HOT BEACH, OLYMPIA, BRAZIL, 54%, 914, 594, 243, 608
10, PARQUE ACUATICO EL ROLLO, MORELOS, MEXICO, 33%, 489, 368, 210, 525"

top10_wp_la = read.table(text = data_text_wp_la, sep = ",", header = FALSE, quote = "", na.strings = "-")
colnames(top10_wp_la) = c("id","Park_Name", "City", "Country", "%Change", "2022", "2021", "2020", "2019")
```


### Clean the dataset for the top 10 water parks Latin America in 2019 to 2022
```{r, echo=TRUE}
top10_wp_la_clean_df =
  top10_wp_la |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Water Park",
    Region = "Latin America",
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```



### Import water parks data of Asia Pacific 2019-2022
```{r, echo=FALSE}
data_text_wp_apac = "1, CHIMELONG WATER PARK, GUANGZHOU, CHINA, -27%, 1620, 2230, 1512, 3014
2, WUHU FANTAWILD WATER PARK, WUHU, CHINA, 115%, 928, 431, 770, 1348
3, CARIBBEAN BAY, SEOUL, SOUTH KOREA, 348%, 995, 222, 168, 1333
4, PLAYA MAYA WATER PARK, WUHAN, CHINA, -23%, 560, 730, 930, 1210
5, AQUAVENTURE ATLANTIS, SANYA, CHINA, -48%, 630, 1200, 1200, 1200
6, SUNWAY LAGOON, KUALA LUMPUR, MALAYSIA, 58%, 950, 600, 600, 1200
7, WET 'N' WILD GOLD COAST, GOLD COAST, AUSTRALIA, 6%, 1030, 970, 952, 1120
8, OCEAN WORLD, HONGCHEON, SOUTH KOREA, 552%, 750, 115, 180, 1071
9, POSEIDON BEACH WATER WORLD, HARBIN, CHINA, 7%, 625, 582, 300, 917
10, SUNWAY LOST WORLD OF TAMBUN, PERAK, MALAYSIA, 33%, 800, 600, 600, 1000
11, PLAYA MAYA WATER PARK, SHANGHAI, CHINA, -73%, 290, 1080, 860, 970
12, ATLANTIS WATER ADVENTURE, JAKARTA, INDONESIA, 777%, 263, 30, 220, 910
13, ZHENGZHOU FANTAWILD WATER PARK, ZHENGZHOU, CHINA, 9%, 484, 443, 622, 905
14, SUMMERLAND, TOKYO, JAPAN, 6%, 690, 650, 454, 868
15, XIAMEN FANTAWILD WATERPARK, XIAMEN, CHINA, 23%, 674, 548, 442, 832
16, THE JUNGLE WATER ADVENTURE, BOGOR, INDONESIA, 117%, 582, 268, 200, 785
17, PLAYA MAYA WATER PARK, TIANJIN, CHINA, -14%, 630, 730, 550, 750
18, WOONGJIN PLAYDOCI WATERDOCI, BUCHEON, SOUTH KOREA, 252%, 327, 93, 146, 750
19, LOTTE WATER PARK, GIMHAE, SOUTH KOREA, 603%, 548, 78, 133, 740
20, ADVENTURE COVE WATER PARK, Sentosa, SINGAPORE, 9%, 360, 330, 300, 710"

top20_wp_apac = read.table(text = data_text_wp_apac, sep = ",", header = FALSE, quote = "", na.strings = "-")
colnames(top20_wp_apac) = c("id","Park_Name", "City", "Country", "%Change", "2022", "2021", "2020", "2019")
```


### Clean the dataset for the top 20 water parks Asia Pacific in 2019 to 2022
```{r, echo=TRUE}
top20_wp_apac_clean_df =
  top20_wp_apac |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Water Park",
    Region = "Asia Pacific",
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```


### Import the water parks data of EMEA
```{r, echo=FALSE}
data_text_wp_emea = "1, THERME ERDING, ERDING, GERMANY, 143%, 1700, 700, 750, 1850
2, AQUAVENTURE WATER PARK, DUBAI, U.A.E., 88%, 1500, 800, 600, 1322
3, AQUAPALACE, PRAGUE, CZECH REPUBLIC, 107%, 1200, 580, 537, 1300
4, TROPICAL ISLANDS, KRAUSNICK, GERMANY, 130%, 1150, 500, 493, 1233
5, SIAM PARK, SANTA CRUZ DE TENERIFE, SPAIN, 125%, 1150, 511, 97, 1200
6, AQUALAND MORAVIA, PASOHLÁVKY, CZECH REPUBLIC, 88%, 727, 386, 368, 806
7, TIKI POOL, DUINRELL, NETHERLANDS, 88%, 750, 400, 390, 800
8, NETTEBAD, OSNABRÜCK, GERMANY, 87%, 650, 347, 311, 756
9, WILD WADI, DUBAI, U.A.E., 60%, 750, 470, 350, 740
10, LALANDIA, BILLUND, DENMARK, 67%, 750, 450, 240, 682"
top10_wp_emea = read.table(text = data_text_wp_emea, sep = ",", header = FALSE, quote = "")
colnames(top10_wp_emea) = c("id","Park_Name", "City", "Country", "%Change", "2022", "2021", "2020", "2019")
```


### Clean the dataset for the top 10 water parks Europe Middle East Africa in 2019 to 2022
```{r, echo=TRUE}
top10_wp_emea_clean_df =
  top10_wp_emea |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Water Park",
    Region = "EMEA",
    Country_State = Country,
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```


### Import museums data of Worldwide 2019-2022
```{r, echo=FALSE}
data_text_mu_ww = "1, LOUVRE, PARIS, FRANCE, 173%, 7726, 2825, 2700, 9600
2, NATIONAL MUSEUM OF CHINA, BEIJING, CHINA, -31%, 1631, 2378, 1600, 7390
3, VATICAN MUSEUMS, VATICAN, VATICAN CITY, 215%, 5081, 1613, 1300, 6883
4, THE METROPOLITAN MUSEUM OF ART, NEW YORK, U.S., 64%, 3209, 1958, 1125, 6770
5, BRITISH MUSEUM, LONDON, U.K., 209%, 4097, 1327, 1275, 6208
6, TATE MODERN, LONDON, U.K., 236%, 3883, 1156, 1433, 6098
7, NATIONAL GALLERY, LONDON, U.K., 285%, 2727, 709, 1197, 6011
8, NATURAL HISTORY MUSEUM, LONDON, U.K., 198%, 4655, 1564, 1197, 5424
9, AMERICAN MUSEUM OF NATURAL HISTORY, NEW YORK, U.S., 0, 0, 0, 0, 5000
10, STATE HERMITAGE, ST. PETERSBURG, RUSSIA, 71%, 2813, 1649, 969, 4957
11, SHANGHAI SCIENCE AND TECHNOLOGY MUSEUM, SHANGHAI, CHINA, -79%, 494, 2368, 1351, 4824
12, REINA SOFÍA, MADRID, SPAIN, 84%, 3063, 1643, 1248, 4426
13, NATIONAL MUSEUM OF NATURAL HISTORY, WASHINGTON, U.S., 297%, 3900, 982, 573, 4200
14, NANJING MUSEUM, NANJING, CHINA, -21%, 1610, 2034, 1515, 4169
15, ZHEJIANG MUSEUM, HANGZHOU, CHINA, -6%, 874, 929, 1108, 4150
16, NATIONAL GALLERY OF ART, WASHINGTON, U.S., 91%, 3256, 1705, 730, 4074
17, VICTORIA & ALBERT MUSEUM, LONDON, U.K., 176%, 2370, 858, 872, 3921
18, CHINA SCIENCE TECHNOLOGY MUSEUM, BEIJING, CHINA, -39%, 1431, 2360, 690, 3891
19, NATIONAL PALACE MUSEUM, TAIPEI, TAIWAN, 149%, 1040, 417, 643, 3832
20, MUSÉE D'ORSAY, PARIS, FRANCE, 213%, 3270, 1044, 867, 3652"

top20_mu_ww = read.table(text = data_text_mu_ww, sep = ",", header = FALSE, quote = "", na.strings = "-")
colnames(top20_mu_ww) = c("id","Park_Name", "City", "Country", "%Change", "2022", "2021", "2020", "2019")
```


### Clean the dataset for the top 20 museums Worldwide in 2019 to 2022
```{r, echo=TRUE}
top20_mu_ww_clean_df =
  top20_mu_ww |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Museum",
    Region = "Worldwide",
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```


### Import museums data of North America 2019-2022
```{r, echo=FALSE}
data_text_mu_na = "1, THE METROPOLITAN MUSEUM OF ART, NEW YORK, NY, U.S., 64%, 3209, 1958, 1125, 6770
2, AMERICAN MUSEUM OF NATURAL HISTORY, NEW YORK, NY, U.S., 0, 0, 0, 0, 5000
3, NATIONAL MUSEUM OF NATURAL HISTORY, WASHINGTON, DC, U.S., 297%, 3900, 982, 573, 4200
4, NATIONAL GALLERY OF ART, WASHINGTON, DC, U.S., 91%, 3256, 1705, 730, 4074
5, NATIONAL AIR AND SPACE MUSEUM, WASHINGTON, DC, U.S., 2%, 419, 411, 267, 3200
6, NATIONAL MUSEUM OF AMERICAN HISTORY, WASHINGTON, DC, U.S., 216%, 1800, 570, 311, 2800
7, CALIFORNIA SCIENCE CENTER, LOS ANGELES, CA, U.S., 133%, 1694, 726, 271, 2200
8, NATIONAL MUSEUM OF AFRICAN AMERICAN HISTORY AND CULTURE, WASHINGTON, DC, U.S., 178%, 1093, 393, 368, 2000
9, SMITHSONIAN AMERICAN ART MUSEUM (WITH THE RENWICK GALLERY), WASHINGTON, DC, U.S., 162%, 1100, 420, 387, 2000
10, THE MUSEUM OF MODERN ART, NEW YORK, NY, U.S., 89%, 2190, 1161, 706, 1992
11, HOUSTON MUSEUM OF NATURAL SCIENCE, HOUSTON, TX, U.S., 40%, 1520, 1086, 0, 1974
12, NATIONAL PORTRAIT GALLERY, WASHINGTON, DC, U.S., 189%, 954, 330, 321, 1700
13, THE ART INSTITUTE OF CHICAGO, CHICAGO, IL, U.S., 38%, 1037, 700, 375, 1666
14, U.S. HOLOCAUST MEMORIAL MUSEUM, WASHINGTON, DC, U.S., 164%, 1184, 448, 0, 1600
15, FIELD MUSEUM OF NATURAL HISTORY, CHICAGO, IL, U.S., 68%, 1018, 605, 316, 1494
16, DENVER MUSEUM OF NATURE & SCIENCE, DENVER, CO, U.S., 20%, 1151, 959, 991, 1487
17, MUSEUM OF SCIENCE, BOSTON, MA, U.S., 93%, 1123, 583, 345, 1458
18, THE J. PAUL GETTY CENTER, LOS ANGELES, CA, U.S., 80%, 909, 504, 267, 1439
19, MUSEUM OF SCIENCE AND INDUSTRY, CHICAGO, IL, U.S., 45%, 1134, 780, 291, 1388
20, CALIFORNIA ACADEMY OF SCIENCES, SAN FRANCISCO, CA, U.S., 50%, 966, 646, 240, 1321"

top20_mu_na = read.table(text = data_text_mu_na, sep = ",", header = FALSE, quote = "", na.strings = "-")
colnames(top20_mu_na) = c("id","Park_Name", "City", "State", "Country", "%Change", "2022", "2021", "2020", "2019")
```


### Clean the dataset for the top 20 museums North America in 2019 to 2022
```{r, echo=TRUE}
top20_mu_na_clean_df =
  top20_mu_na |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Museum",
    Region = "North America",
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```


### Import museums data of Asia Pacific 2019-2022
```{r, echo=FALSE}
data_text_mu_apac = "1, NATIONAL MUSEUM OF CHINA, BEIJING, CHINA, -31%, 1631, 2378, 1600, 7390
2, SHANGHAI SCIENCE & TECHNOLOGY MUSEUM, SHANGHAI, CHINA, -79%, 494, 2368, 1351, 4824
3, NANJING MUSEUM, NANJING, CHINA, -21%, 1610, 2034, 1515, 4169
4, ZHEJIANG MUSEUM, HANGZHOU, CHINA, -6%, 874, 929, 1108, 4150
5, CHINA SCIENCE TECHNOLOGY MUSEUM, BEIJING, CHINA, -39%, 1431, 2360, 690, 3891
6, NATIONAL PALACE MUSEUM, TAIPEI, TAIWAN, 149%, 1040, 417, 643, 3832
7, XIAN MUSEUM, XIAN, CHINA, 27%, 401, 315, 348, 3420
8, NATIONAL MUSEUM OF KOREA, SEOUL, SOUTH KOREA, 170%, 3411, 1263, 774, 3354
9, TIANJIN NATURAL HISTORY MUSEUM, TIANJIN, CHINA, -17%, 1321, 1600, 550, 3150
10, HUNAN MUSEUM, CHANGSHA, CHINA, -14%, 1221, 1421, 1130, 3140
11, NATIONAL TAIWAN SCIENCE EDUCATION CENTER, TAIPEI, TAIWAN, 62%, 1879, 1158, 1082, 3000
12, SHANXI HISTORY MUSEUM, XIAN, CHINA, -54%, 690, 1510, 1000, 2900
13, TOKYO METROPOLITAN ART MUSEUM, TOKYO, JAPAN, 44%, 1510, 1049, 510, 2874
14, NATIONAL MUSEUM OF NATURAL SCIENCE, TAICHUNG, TAIWAN, 47%, 2300, 1566, 2263, 2650
15, SHANGHAI NATURAL HISTORY MUSEUM, SHANGHAI, CHINA, -75%, 532, 2166, 1040, 2549
16, NATIONAL MUSEUM OF NATURE AND SCIENCE, TOKYO, JAPAN, 69%, 1350, 800, 580, 2460
17, ZHEJIANG MUSEUM OF NATURAL HISTORY, HANGZHOU, CHINA, 0%, 738, 740, 600, 2443
18, NATIONAL GALLERY OF VICTORIA, MELBOURNE, AUSTRALIA, 97%, 1580, 802, 637, 2433
19, TOKYO NATIONAL MUSEUM, TOKYO, JAPAN, 127%, 1372, 605, 725, 2431
20, HUBEI PROVINCIAL MUSEUM, WUHAN, CHINA, 37%, 1000, 730, 446, 2400"

top20_mu_apac = read.table(text = data_text_mu_apac, sep = ",", header = FALSE, quote = "", na.strings = "-")
colnames(top20_mu_apac) = c("id","Park_Name", "City", "Country", "%Change", "2022", "2021", "2020", "2019")
```


### Clean the dataset for the top 20 museums Asia Pacific in 2019 to 2022
```{r, echo=TRUE}
top20_mu_apac_clean_df =
  top20_mu_apac |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Museum",
    Region = "Asia Pacific",
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```


### Import museums data of Europe Middle East Africa 2019-2022
```{r, echo=FALSE}
data_text_mu_emea = "1, LOUVRE, PARIS, FRANCE, 173%, 7726, 2825, 2700, 9600
2, VATICAN MUSEUMS, VATICAN, VATICAN CITY, 215%, 5081, 1613, 1300, 6883
3, BRITISH MUSEUM, LONDON, U.K., 209%, 4097, 1327, 1275, 6208
4, TATE MODERN, LONDON, U.K., 236%, 3883, 1156, 1433, 6098
5, NATIONAL GALLERY, LONDON, U.K., 285%, 2727, 709, 1197, 6011
6, NATURAL HISTORY MUSEUM, LONDON, U.K., 198%, 4655, 1564, 1197, 5424
7, STATE HERMITAGE, ST PETERSBURG, RUSSIA, 71%, 2813, 1649, 969, 4957
8, REINA SOFÍA, MADRID, SPAIN, 86%, 3063, 1643, 1248, 4426
9, VICTORIA & ALBERT MUSEUM, LONDON, U.K., 176%, 2370, 858, 872, 3921
10, MUSÉE D'ORSAY, PARIS, FRANCE, 213%, 3270, 1044, 867, 3652
11, CENTRE POMPIDOU, PARIS, FRANCE, 101%, 3010, 1501, 913, 3270
12, SCIENCE MUSEUM (SOUTH KENSINGTON), LONDON, U.K., 262%, 2335, 646, 862, 3254
13, MUSEO NACIONAL DEL PRADO, MADRID, SPAIN, 118%, 2457, 1127, 852, 3203
14, STATE TRETYAKOV GALLERY, MOSCOW, RUSSIA, 21%, 1910, 1581, 894, 2836
15, RIJKSMUSEUM, AMSTERDAM, NETHERLANDS, 183%, 1766, 625, 675, 2700
16, CITE DES SCIENCES ET DE L'INDUSTRIE, PARIS, FRANCE, 114%, 2101, 984, 793, 2370
17, GALLERIA DEGLI UFFIZI, FLORENCE, ITALY, 129%, 2223, 970, 659, 2362
18, AUSCHWITZ-BIRKENAU MUSEUM, OŚWIĘCIM, POLAND, 110%, 1184, 563, 502, 2300
19, NATIONAL MUSEUM OF SCOTLAND, EDINBURGH, U.K., 199%, 1974, 661, 444, 2210
20, VAN GOGH MUSEUM, AMSTERDAM, NETHERLANDS, 267%, 1346, 366, 517, 2100"

top20_mu_emea = read.table(text = data_text_mu_emea, sep = ",", header = FALSE, quote = "", na.strings = "-")
colnames(top20_mu_emea) = c("id","Park_Name", "City", "Country", "%Change", "2022", "2021", "2020", "2019")
```


### Clean the dataset for the top 20 museums Europe Middle East Africa in 2019 to 2022
```{r, echo=TRUE}
top20_mu_emea_clean_df =
  top20_mu_emea |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Museum",
    Region = "Europe Middle East Africa",
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```




### Combine all dataset
```{r, echo=TRUE}
combine_df = 
  bind_rows(top25_tp_ww_clean_df, top20_tp_na_clean_df, top10_tp_la_clean_df, top20_tp_apac_clean_df, top20_tp_emea_clean_df, top20_wp_na_clean_df, top10_wp_la_clean_df, top20_wp_apac_clean_df, top10_wp_emea_clean_df, top20_mu_ww_clean_df, top20_mu_na_clean_df, top20_mu_apac_clean_df, top20_mu_emea_clean_df)
```


### Export the final dataset
```{r, echo=TRUE}
write.csv(combine_df, "ultimate data.csv", row.names = FALSE)
```





