---
title: "Exploratory Analysis"
---

```{r setup, include=FALSE}
library(tidyverse)
library(plotly)
library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

After tidying our data set, an exploratory analysis is conducted to look for possible predictors for the `Attendance` outcome. 

```{r, include=FALSE}
theme_park = read_csv("ultimate data.csv") |> 
    mutate(
    Year = as.factor(Year),
    Type = as.factor(Type)
  ) |> 
    filter(
      Region != c("Worldwide")
    ) 
```

## Descriptive Analysis for All Data: 

### Total Attendance by Type from 2019 to 2022

A brief summary of attendance based on the `Type` variable is provided below:

```{r region attend summary}
theme_park  |> 
    group_by(Year, Type) |> 
    mutate(
      Attendance = Attendance / 100000
    ) |> 
    summarise(sum = sum(Attendance)) |> 
    arrange(Type) |> 
    pivot_wider(
      names_from = Type,
      values_from = sum
    ) |> 
    knitr::kable(digits = 3, caption = c("Summary of Attendance for Three Types of Facilities From 2019 to 2022"))
```

From this table, some observed patterns are:

* A common decreasing trend of attendance from 2019 and 2020
* A common increasing trend of attendance from 2020 and 2022

The distribution of data by year is further visualized into the box plots below:

```{r}
theme_park |> 
  group_by(Year) |> 
  plot_ly(y = ~Attendance, color = ~Year, type = "box", colors = "viridis")
```

* Overall high attendance in 2019 and low attendance in 2020
* Common outliers in the distribution of attendance for each year

### Total Attendance by Region from 2019 to 2022

Next, we specifically look at the trend of `Attendance` from 2019 to 2022 based on the `Region` variable. 

```{r}
 theme_park|> 
  group_by(Region, Year)  |> 
  summarize(attend_sum = mean(Attendance)) |> 
  plot_ly(x = ~Year, y = ~attend_sum, color = ~Region,
          type = "scatter", mode = 'point', colors = "viridis")
```

* Similar attendance fluctuation for most regions
* Only EMEA has an attendance drop from 2021 to 2022

## ANOVA TEST

### Based on Type of Facilities

The first ANOVA test focuses on the `Type` variable in our data set. The null hypothesis and alternative hypothesis are listed as follow:

$$H_0: \mu_{\text{Amusement/Theme Park}} = \mu_{\text{Water Park}} = \mu_{\text{Museum}} ~~ \text{vs} ~~ H_1: \text{at least two means are not equal}$$

```{r}

anova_1 = aov(Attendance ~ Type, data = theme_park)

summary(anova_1)

```

With a p-value of less than 2e-16, we would reject the null hypothesis. We have evidence that at least
two of the means are not equal. Meaning the mean attendance among type groups is different
for at least two groups in the `Type` variable.

