---
title: "Impact of the Pandemic on the Attendance of Theme Parks/Museums Worldwide"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

## Motivation

As theme parks and museums are significant contributors to the tourism and entertainment industry. Then understanding how attendance has been affected is vital and may provides insights into the economic impact of the pandemic on these sectors. \newline

In addition, analyzing attendance data of parks and museums from different region can provide insights into how the impact of the pandemic on theme parks and museums in the United States compares to other countries. This comparative analysis helps in understanding the global dynamics of the tourism and entertainment industries, which also provides insights into understanding how different regions responded to COVID-19.

## Related work

The official websites of major theme park operators release statements and reports on how the pandemic has affected their attendance and operations, which inspire us to investigate the impact of the COVID-19 pandemic on theme park attendance.

## Initial questions

The main question we are trying to understand is what is the overall  Impact of the COVID-19 Pandemic on Theme Park Attendance. While collecting the data, we also involved attendance data of different theme parks, water parks and museums around the world. After adding variables such as park types and regions, we also tried to investigate the relationship between the park attendance and other covariants.

## Data Collection and cleaning

```{r include=FALSE}
library(tidyverse)
library(dplyr)
library(readxl)
library(openxlsx)
```

### Data Collection and Data Source
We sourced our data from the *"Global Attractions Attendance Report"*, a publication by AECOM and TEA. The dataset comprises attendance information for various theme parks and museums worldwide, covering the years 2019 to 2022. You can access the detailed report through this [link](https://aecom.com/wp-content/uploads/documents/reports/AECOM-Theme-Index-2022.pdf).

### Data Import and Cleaning
Each individual dataset catgegorized by types of parks and regions of the parks are listed in the article. By collecting each of the dataset, we created separate dataframes and combine to one ultimate dataset with all of the data focusing on `Park_Name`, `City`, `Country`, `Year`, `Attendance`, `Type`, and `Region`.

* `Park_Name`: Names of the park
* `City`: City of where the park is located
* `Country`: Country of where the park is located
* `Year`: Year of attendance
* `Attendance`: Number of people visited the park
* `Type`: Amusement/Theme Parks, Water Parks, and Museums
* `Region`: Worldwide, North America, Latin America, Asia Pacific, Europe Middle East Africa


#### Import the amusement/theme parks data of Worldwide 2019-2022
```{r, echo=TRUE}
data_text_tp_ww = "1, MAGIC KINGDOM THEME PARK AT WALT DISNEY WORLD RESORT, LAKE BUENA VISTA, U.S., 35%, 17133, 12691, 6941, 20963
2, DISNEYLAND PARK AT DISNEYLAND RESORT, ANAHEIM, U.S., 97%, 16881, 8573, 3674, 18666
3, TOKYO DISNEYLAND AT TOKYO DISNEY RESORT, TOKYO, JAPAN, 90%, 12000, 6300, 4160, 17910
4, TOKYO DISNEYSEA AT TOKYO DISNEY RESORT, TOKYO, JAPAN, 74%, 10100, 5800, 3400, 14650
5, UNIVERSAL STUDIOS JAPAN, OSAKA, JAPAN, 125%, 12350, 5500, 4901, 14500
6, DISNEY'S ANIMAL KINGDOM AT WALT DISNEY WORLD, LAKE BUENA VISTA, U.S., 25%, 9027, 7194, 4166, 13888
7, EPCOT AT WALT DISNEY WORLD, LAKE BUENA VISTA, U.S., 29%, 10000, 7752, 4044, 12444
8, CHIMELONG OCEAN KINGDOM, HENGQIN, CHINA, -41%, 4400, 7452, 4797, 11736
9, DISNEY'S HOLLYWOOD STUDIOS AT WALT DISNEY WORLD, LAKE BUENA VISTA, U.S., 27%, 10900, 8589, 3675, 11483
10, SHANGHAI DISNEYLAND AT SHANGHAI DISNEY RESORT, SHANGHAI, CHINA, -38%, 5300, 8480, 5500, 11210
11, UNIVERSAL STUDIOS FLORIDA AT UNIVERSAL ORLANDO, FL, U.S., 20%, 10750, 8987, 4096, 10922
12, UNIVERSAL'S ISLANDS OF ADVENTURE AT UNIVERSAL ORLANDO, ORLANDO, U.S., 21%, 11025, 9077, 4005, 10375
13, DISNEYCALIFORNIA ADVENTURE PARK AT DISNEYLAND RESORT, ANAHEIM, U.S., 81%, 9000, 4977, 1919, 9861
14, DISNEYLAND PARK AT DISNEYLAND PARIS, MARNE-LA-VALLEE, FRANCE, 184%, 9930, 3500, 2620, 9745
15, UNIVERSAL STUDIOS HOLLYWOOD, UNIVERSAL CITY, U.S., 53%, 8400, 5505, 1299, 9147
16, EVERLAND, GYEONGGI-DO, SOUTH KOREA, 56%, 5770, 3710, 2760, 6606
17, LOTTE WORLD, SEOUL, SOUTH KOREA, 84%, 4520, 2460, 1560, 5953
18, NAGASHIMA SPA LAND, KUWANA, JAPAN, 17%, 4200, 3600, 2400, 5950
19, EUROPA-PARK, RUST, GERMANY, 100%, 6000, 3000, 2500, 5750
20, OCEAN PARK, HONG KONG, SAR, 0%, 1400, 1400, 2200, 5700
21, HONG KONG DISNEYLAND, HONG KONG, SAR, 21%, 3400, 2800, 1700, 5695
22, DE EFTELING EFTELING THEME PARK RESORT, KAATSHEUVEL, NETHERLANDS, 65%, 5430, 3300, 2900, 5260
23, WALT DISNEY STUDIOS PARK AT DISNEYLAND PARIS, MARNE-LA-VALLEE, FRANCE, 183%, 5340, 1884, 1410, 5245
24, OCT HAPPY VALLEY, BEIJING, CHINA, -24%, 3740, 4930, 3950, 5160
25, CHIMELONG PARADISE, GUANGZHOU, CHINA, -41%, 2300, 3890, 2681, 4905"

top25_tp_ww = read.table(text = data_text_tp_ww, sep = ",", header = FALSE, quote = "", na.strings = "-")
colnames(top25_tp_ww) = c("id","Park_Name", "City", "Country", "%Change", "2022", "2021", "2020", "2019")
```


#### Clean the dataset for the top 25 amusement/theme parks Worldwide in 2019 to 2022
```{r, echo=TRUE}
top25_tp_ww_clean_df =
  top25_tp_ww |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Amusement/Theme Park",
    Region = "Worldwide",
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```


#### Import the amusement/theme parks data of North America 2019-2022
```{r, echo=TRUE}
data_text_tp_na = "1, MAGIC KINGDOM THEME PARK AT WALT DISNEY WORLD, LAKE BUENA VISTA, FL, U.S., 35%, 17133, 12691, 6941, 20963
2, DISNEYLAND PARK AT DISNEYLAND RESORT, ANAHEIM, CA, U.S., 97%, 16881, 8573, 3674, 18666
3, DISNEY'S ANIMAL KINGDOM AT WALT DISNEY WORLD, LAKE BUENA VISTA, FL, U.S., 25%, 9027, 7194, 4166, 13888
4, EPCOT AT WALT DISNEY WORLD, LAKE BUENA VISTA, FL, U.S., 29%, 10000, 7752, 4044, 12444
5, DISNEY'S HOLLYWOOD STUDIOS AT WALT DISNEY WORLD, LAKE BUENA VISTA, FL, U.S., 27%, 10900, 8589, 3675, 11483
6, UNIVERSAL STUDIOS FLORIDA, ORLANDO, FL, U.S., 20%, 10750, 8987, 4096, 10922
7, UNIVERSAL'S ISLANDS OF ADVENTURE, ORLANDO, FL, U.S., 21%, 11025, 9077, 4005, 10375
8, DISNEY CALIFORNIA ADVENTURE PARK AT DISNEYLAND RESORT, ANAHEIM, CA, U.S., 81%, 9000, 4977, 1919, 9861
9, UNIVERSAL STUDIOS HOLLYWOOD, UNIVERSAL CITY, CA, U.S., 53%, 8400, 5505, 1299, 9147
10, SEAWORLD ORLANDO, ORLANDO, FL, U.S., 46%, 4454, 3051, 1598, 4640
11, KNOTT'S BERRY FARM, BUENA PARK, CA, U.S., 6%, 3899, 3681, 811, 4238
12, BUSCH GARDENS TAMPA BAY, TAMPA, FL, U.S., 26%, 4051, 3210, 1288, 4180
13, CANADA'S WONDERLAND, MAPLE, ONTARIO, CANADA, 524%, 3768, 587, 0, 3950
14, SEAWORLD SAN DIEGO, SAN DIEGO, CA, U.S., 25%, 3507, 2800, 1139, 3731
15, CEDAR POINT, SANDUSKY, OH, U.S., 4%, 3444, 3327, 1020, 3610
16, SIX FLAGS MAGIC MOUNTAIN, VALENCIA, CA, U.S., -2%, 2993, 3047, 686, 3521
17, KINGS ISLAND, KINGS ISLAND, OH, U.S., 5%, 3340, 3181, 1626, 3485
18, SIX FLAGS GREAT ADVENTURE, JACKSON, NJ, U.S., -26%, 2153, 2913, 598, 3451
19, HERSHEYPARK, HERSHEY, PA, U.S., 6%, 3193, 3012, 1717, 3384
20, SIX FLAGS GREAT AMERICA, GURNEE, IL, U.S., -5%, 2535, 2675, 0, 3169"

top20_tp_na = read.table(text = data_text_tp_na, sep = ",", header = FALSE, quote = "")
colnames(top20_tp_na) = c("id","Park_Name", "City", "State", "Country", "%Change", "2022", "2021", "2020", "2019")
```


#### Clean the dataset for the top 20 amusement/theme parks North America in 2019 to 2022
```{r, echo=TRUE}
top20_tp_na_clean_df =
  top20_tp_na |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Amusement/Theme Park",
    Region = "North America",
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```


#### Import the amusement/theme parks data of Latin America 2019-2022
```{r, echo=TRUE}
data_text_tp_la = "1,SIX FLAGS MEXICO, MEXICO CITY, MEXICO, 54%, 1727, 1125, 701, 2803
2,BETO CARRERO WORLD, SANTA CATARINA, BRAZIL, 23%, 2327, 1895, 1252, 2241
3,PARQUE XCARET, CANCUN, MEXICO, 27%, 1399, 1104, 736, 1960
4,MUNDO PETAPA, GUATEMALA CITY, GUATEMALA, 99%, 800, 402, 268, 1330
5,LA FERIA DE CHAPULTEPEC, MEXICO CITY, MEXICO, 0%, 0, 0, 0, 1306
6,PLAZA DE SESAMO, MONTERREY, MEXICO, 0%, 0, 0, 0, 1173
7,PARQUE MUNDO AVENTURA, BOGOTÁ, COLOMBIA, 19%, 752, 631, 344, 1151
8,FANTASIALANDIA, SANTIAGO, CHILE, 38%, 850, 615, 430, 1100
9,THEME PARQUE NACIONAL DEL CAFÉ, QUINDIO, COLOMBIA, 30%, 626, 482, 321, 1043
10,PARQUE DE LA COSTA, BUENOS AIRES, ARGENTINA, 30%, 514, 395, 263, 968"

top10_tp_la = read.table(text = data_text_tp_la, sep = ",", header = FALSE, quote = "")
colnames(top10_tp_la) = c("id","Park_Name", "City", "Country", "%Change", "2022", "2021", "2020", "2019")
```


#### Clean the dataset for the top 10 amusement/theme parks Latin America in 2019 to 2022
```{r, echo=TRUE}
top10_tp_la_clean_df =
  top10_tp_la |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Amusement/Theme Park",
    Region = "Latin America",
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```


#### Import the amusement/theme parks data of Asia Pacific 2019-2022
```{r, echo=TRUE}
data_text_tp_apac = "1, TOKYO DISNEYLAND AT TOKYO DISNEY RESORT, TOKYO, JAPAN, 90%, 12000, 6300, 4160, 17910
2, TOKYO DISNEYSEA AT TOKYO DISNEY RESORT, TOKYO, JAPAN, 74%, 10100, 5800, 3400, 14650
3, UNIVERSAL STUDIOS JAPAN, OSAKA, JAPAN, 125%, 12350, 5500, 4901, 14500
4, CHIMELONG OCEAN KINGDOM, HENGQIN, CHINA, -41%, 4400, 7452, 4797, 11736
5, SHANGHAI DISNEYLAND AT SHANGHAI DISNEY RESORT, SHANGHAI, CHINA, -38%, 5300, 8480, 5500, 11210
6, EVERLAND, SEOUL, SOUTH KOREA, 56%, 5770, 3710, 2760, 6606
7, LOTTE WORLD, SEOUL, SOUTH KOREA, 84%, 4520, 2460, 1560, 5953
8, NAGASHIMA SPA LAND, KUWANA, JAPAN, 17%, 4200, 3600, 2400, 5950
9, OCEAN PARK, HONG KONG, SAR, 0%, 1400, 1400, 2200, 5700
10, HONG KONG DISNEYLAND, HONG KONG, SAR, 21%, 3400, 2800, 1700, 5695
11, OCT HAPPY VALLEY, BEIJING, CHINA, -24%, 3740, 4930, 3950, 5160
12, CHIMELONG PARADISE, GUANGZHOU, CHINA, -41%, 2300, 3890, 2681, 4905
13, UNIVERSAL STUDIOS SINGAPORE, Sentosa, SINGAPORE, 75%, 2100, 1200, 1098, 4500
14, CHINA DINOSAUR PARK, CHANGZHOU, CHINA, -12%, 2200, 2500, 2375, 4434
15, OCT WINDOW OF THE WORLD, SHENZHEN, CHINA, -54%, 1100, 2400, 1890, 3990
16, OCT HAPPY VALLEY, SHENZHEN, CHINA, -47%, 1750, 3310, 3120, 3980
17, ZHENGZHOU FANTAWILD ADVENTURE, ZHENGZHOU, CHINA, -42%, 1590, 2752, 3421, 3840
18, OCT HAPPY VALLEY, CHENGDU, CHINA, -47%, 1570, 2940, 2620, 3580
19, NINGBO FANTAWILD ORIENTAL HERITAGE, NINGBO, CHINA, -47%, 1230, 2308, 2257, 3575
20, OCT HAPPY VALLEY, SHANGHAI, CHINA, -62%, 1240, 3300, 2730, 3390"

top20_tp_apac = read.table(text = data_text_tp_apac, sep = ",", header = FALSE, quote = "", na.strings = "-")
colnames(top20_tp_apac) = c("id","Park_Name", "City", "Country", "%Change", "2022", "2021", "2020", "2019")
```


#### Clean the dataset of the top 20 Amusement/Theme Parks Asia Pacific in 2019 to 2022
```{r, echo=TRUE}
top20_tp_apac_clean_df =
  top20_tp_apac |>
  pivot_longer(
    ("2019"):("2021"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Amusement/Theme Park",
    Region = "Asia Pacific",
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```


#### Import the amusement/theme parks data of Europe Middle East Africa
```{r, echo=TRUE}
data_text_tp_emea = "1,DISNEYLAND PARK AT DISNEYLAND PARIS, MARNE-LA-VALLEE, FRANCE,184%,9930,3500,2620,9745
2,EUROPA-PARK, RUST, GERMANY,80%,5400,3000,2500,5750
3,DE EFTELING THEME PARK RESORT, KAATSHEUVEL, NETHERLANDS,65%,5430,3300,2900,5260
4,WALT DISNEY STUDIOS PARK AT DISNEYLAND PARIS, MARNE-LA-VALLEE, FRANCE,183%,5340,1884,1410,5245
5,TIVOLI GARDENS, COPENHAGEN, DENMARK,61%,3864,2400,1628,4581
6,PORTAVENTURA WORLD, SALOU, SPAIN,56%,3750,2400,700,3750
7,LISEBERG, GOTHENBURG, SWEDEN,73%,2500,1447,0,2950
8,GARDALAND, CASTELNUOVO DEL GARDA, ITALY,34%,2950,2200,1350,2920
9,LEGOLAND WINDSOR, WINDSOR, U.K.,60%,2400,1500,450,2430
10,PARC ASTERIX, PLAILLY, FRANCE,102%,2632,1300,1163,2326
11,PUY DU FOU, LES EPESSES, FRANCE,45%,2342,1616,923,2308
12,PARQUE WARNER, MADRID, SPAIN,43%,1860,1300,450,2232
13,ALTON TOWERS, STAFFORDSHIRE, U.K.,28%,2300,1800,670,2130
14,PHANTASIALAND, BRÜHL, GERMANY,78%,2100,1180,1000,2050
15,LEGOLAND BILLUND, BILLUND, DENMARK,164%,2242,850,700,1950
16,THORPE PARK, CHERTSEY, U.K.,-6%,1600,1700,600,1900
17,FUTUROSCOPE, JAUNAY-CLAN, FRANCE,75%,1920,1100,900,1900
18,LEGOLAND DEUTSCHLAND, GÜNZBURG, GERMANY,89%,1700,900,750,1700
19,HEIDE PARK, SOLTAU, GERMANY,23%,1600,1300,950,1700
20,CHESSINGTON WORLD OF ADVENTURES, CHESSINGTON, U.K.,3%,1500,1450,510,1690"

top20_tp_emea = read.table(text = data_text_tp_emea, sep = ",", header = FALSE, quote = "", na.strings = "-")
colnames(top20_tp_emea) = c("id","Park_Name", "City", "Country", "%Change", "2022", "2021", "2020", "2019")
```


#### Clean the dataset for the top 20 amusement/theme parks Europe Middle East Africa in 2019 to 2022
```{r, echo=TRUE}
top20_tp_emea_clean_df =
  top20_tp_emea |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Amusement/Theme Park",
    Region = "EMEA",
    Country_State = Country,
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```


#### Import water parks data of North America 2019-2022
```{r, echo=TRUE}
data_text_wp_na = "1,DISNEY'S TYPHOON LAGOON WATER PARK AT WALT DISNEY WORLD RESORT, ORLANDO, FL, U.S., 0%, 1915, 0, 0, 2248
2,DISNEY'S BLIZZARD BEACH WATER PARK AT WALT DISNEY WORLD RESORT, ORLANDO, FL, U.S., -92%, 101, 1201, 316, 1983
3,UNIVERSAL'S VOLCANO BAY, ORLANDO, FL, U.S., 9%, 1850, 1691, 551, 1811
4,AQUATICA, ORLANDO, FL, U.S., 35%, 1548, 1147, 528, 1533
5,SCHLITTERBAHN, NEW BRAUNFELS, TX, U.S., 7%, 946, 888, 451, 996
6,WATER COUNTRY USA, WILLIAMSBURG, VA, U.S., 1%, 707, 700, 0, 736
7,ADVENTURE ISLAND, TAMPA, FL, U.S., 7%, 636, 593, 125, 656
8,AQUATICA SAN ANTONIO, SAN ANTONIO, TX, U.S., 8%, 631, 586, 263, 651
9,SCHLITTERBAHN, GALVESTON, TX, U.S., 34%, 539, 402, 337, 567
10,SPLISH SPLASH, CALVERTON, NY, U.S., 0, 0, 0, 0, 542
11,SIX FLAGS HURRICANE HARBOR, ARLINGTON, TX, U.S., 3%, 388, 377, 235, 538
12,SIX FLAGS WHITE WATER ATLANTA, MARIETTA, GA, U.S., -5%, 350, 368, 238, 526
13,SIX FLAGS HURRICANE HARBOR, JACKSON, NJ, U.S., -5%, 426, 448, 85, 482
14,TYPHOON TEXAS, KATY, TX, U.S., 7%, 364, 340, 203, 478
15,SIX FLAGS HURRICANE HARBOR SPLASHTOWN, SPRING, TX, U.S., 3%, 343, 333, 0, 475
16,ZOOMBEZI BAY, POWELL, OH, U.S., 7%, 399, 373, 176, 466
17,SIX FLAGS HURRICANE HARBOR PHOENIX, GLENDALE, AZ, U.S., 3%, 318, 309, 0, 441
18,CAMELBEACH, TANNERSVILLE, PA, U.S., 66%, 373, 225, 151, 439
19,DOLLYWOOD'S SPLASH COUNTRY, PIGEON FORGE, TN, U.S., 31%, 368, 281, 188, 433
20,CEDAR POINT SHORES, SANDUSKY, OH, U.S., 1%, 405, 402, 0, 424"

top20_wp_na = read.table(text = data_text_wp_na, sep = ",", header = FALSE, quote = "")
colnames(top20_wp_na) = c("id","Park_Name", "City", "State", "Country", "%Change", "2022", "2021", "2020", "2019")
```


#### Clean the dataset for the top 20 water parks North America in 2019 to 2022
```{r, echo=TRUE}
top20_wp_na_clean_df =
  top20_wp_na |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Water Park",
    Region = "North America",
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```


#### Import water parks data of Latin America 2019-2022
```{r, echo=TRUE}
data_text_wp_la = "1, THERMAS DOS LARANJAIS, OLIMPIA, BRAZIL, 47%, 1729, 1179, 922, 1845
2, BAHAMAS AQUAVENTURE WATER PARK, PARADISE ISLAND (NASSAU NEW PROVIDENCE), BAHAMAS, 38%, 1751, 1269, 494, 1813
3, HOT PARK RIO QUENTE, RIO QUENTE, BRAZIL, 53%, 1430, 933, 734, 1469
4, PARQUE ACUATICO XOCOMIL, SAN MARTÍN ZAPOTITLÁN, GUATEMALA, 33%, 833, 626, 121, 894
5, BEACH PARK, AQUIRAZ, BRAZIL, 56%, 782, 500, 312, 890
6, PISCILAGO, GIRARDOT (BOGOTÁ), COLOMBIA, 34%, 816, 609, 307, 876
7, MAGIC CITY WATER PARK, SUZANO, BRAZIL, 32%, 616, 467, 355, 681
8, THERMAS WATER PARK, SÃO PEDRO, BRAZIL, 67%, 753, 452, 315, 641
9, HOT BEACH, OLYMPIA, BRAZIL, 54%, 914, 594, 243, 608
10, PARQUE ACUATICO EL ROLLO, MORELOS, MEXICO, 33%, 489, 368, 210, 525"

top10_wp_la = read.table(text = data_text_wp_la, sep = ",", header = FALSE, quote = "", na.strings = "-")
colnames(top10_wp_la) = c("id","Park_Name", "City", "Country", "%Change", "2022", "2021", "2020", "2019")
```


#### Clean the dataset for the top 10 water parks Latin America in 2019 to 2022
```{r, echo=TRUE}
top10_wp_la_clean_df =
  top10_wp_la |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Water Park",
    Region = "Latin America",
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```


#### Import water parks data of Asia Pacific 2019-2022
```{r, echo=TRUE}
data_text_wp_apac = "1, CHIMELONG WATER PARK, GUANGZHOU, CHINA, -27%, 1620, 2230, 1512, 3014
2, WUHU FANTAWILD WATER PARK, WUHU, CHINA, 115%, 928, 431, 770, 1348
3, CARIBBEAN BAY, SEOUL, SOUTH KOREA, 348%, 995, 222, 168, 1333
4, PLAYA MAYA WATER PARK, WUHAN, CHINA, -23%, 560, 730, 930, 1210
5, AQUAVENTURE ATLANTIS, SANYA, CHINA, -48%, 630, 1200, 1200, 1200
6, SUNWAY LAGOON, KUALA LUMPUR, MALAYSIA, 58%, 950, 600, 600, 1200
7, WET 'N' WILD GOLD COAST, GOLD COAST, AUSTRALIA, 6%, 1030, 970, 952, 1120
8, OCEAN WORLD, HONGCHEON, SOUTH KOREA, 552%, 750, 115, 180, 1071
9, POSEIDON BEACH WATER WORLD, HARBIN, CHINA, 7%, 625, 582, 300, 917
10, SUNWAY LOST WORLD OF TAMBUN, PERAK, MALAYSIA, 33%, 800, 600, 600, 1000
11, PLAYA MAYA WATER PARK, SHANGHAI, CHINA, -73%, 290, 1080, 860, 970
12, ATLANTIS WATER ADVENTURE, JAKARTA, INDONESIA, 777%, 263, 30, 220, 910
13, ZHENGZHOU FANTAWILD WATER PARK, ZHENGZHOU, CHINA, 9%, 484, 443, 622, 905
14, SUMMERLAND, TOKYO, JAPAN, 6%, 690, 650, 454, 868
15, XIAMEN FANTAWILD WATERPARK, XIAMEN, CHINA, 23%, 674, 548, 442, 832
16, THE JUNGLE WATER ADVENTURE, BOGOR, INDONESIA, 117%, 582, 268, 200, 785
17, PLAYA MAYA WATER PARK, TIANJIN, CHINA, -14%, 630, 730, 550, 750
18, WOONGJIN PLAYDOCI WATERDOCI, BUCHEON, SOUTH KOREA, 252%, 327, 93, 146, 750
19, LOTTE WATER PARK, GIMHAE, SOUTH KOREA, 603%, 548, 78, 133, 740
20, ADVENTURE COVE WATER PARK, Sentosa, SINGAPORE, 9%, 360, 330, 300, 710"

top20_wp_apac = read.table(text = data_text_wp_apac, sep = ",", header = FALSE, quote = "", na.strings = "-")
colnames(top20_wp_apac) = c("id","Park_Name", "City", "Country", "%Change", "2022", "2021", "2020", "2019")
```


#### Clean the dataset for the top 20 water parks Asia Pacific in 2019 to 2022
```{r, echo=TRUE}
top20_wp_apac_clean_df =
  top20_wp_apac |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Water Park",
    Region = "Asia Pacific",
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```


#### Import the water parks data of Europe Middle East Africa
```{r, echo=TRUE}
data_text_wp_emea = "1, THERME ERDING, ERDING, GERMANY, 143%, 1700, 700, 750, 1850
2, AQUAVENTURE WATER PARK, DUBAI, U.A.E., 88%, 1500, 800, 600, 1322
3, AQUAPALACE, PRAGUE, CZECH REPUBLIC, 107%, 1200, 580, 537, 1300
4, TROPICAL ISLANDS, KRAUSNICK, GERMANY, 130%, 1150, 500, 493, 1233
5, SIAM PARK, SANTA CRUZ DE TENERIFE, SPAIN, 125%, 1150, 511, 97, 1200
6, AQUALAND MORAVIA, PASOHLÁVKY, CZECH REPUBLIC, 88%, 727, 386, 368, 806
7, TIKI POOL, DUINRELL, NETHERLANDS, 88%, 750, 400, 390, 800
8, NETTEBAD, OSNABRÜCK, GERMANY, 87%, 650, 347, 311, 756
9, WILD WADI, DUBAI, U.A.E., 60%, 750, 470, 350, 740
10, LALANDIA, BILLUND, DENMARK, 67%, 750, 450, 240, 682"
top10_wp_emea = read.table(text = data_text_wp_emea, sep = ",", header = FALSE, quote = "")
colnames(top10_wp_emea) = c("id","Park_Name", "City", "Country", "%Change", "2022", "2021", "2020", "2019")
```


#### Clean the dataset for the top 10 water parks Europe Middle East Africa in 2019 to 2022
```{r, echo=TRUE}
top10_wp_emea_clean_df =
  top10_wp_emea |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Water Park",
    Region = "EMEA",
    Country_State = Country,
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```


#### Import museums data of Worldwide 2019-2022
```{r, echo=TRUE}
data_text_mu_ww = "1, LOUVRE, PARIS, FRANCE, 173%, 7726, 2825, 2700, 9600
2, NATIONAL MUSEUM OF CHINA, BEIJING, CHINA, -31%, 1631, 2378, 1600, 7390
3, VATICAN MUSEUMS, VATICAN, VATICAN CITY, 215%, 5081, 1613, 1300, 6883
4, THE METROPOLITAN MUSEUM OF ART, NEW YORK, U.S., 64%, 3209, 1958, 1125, 6770
5, BRITISH MUSEUM, LONDON, U.K., 209%, 4097, 1327, 1275, 6208
6, TATE MODERN, LONDON, U.K., 236%, 3883, 1156, 1433, 6098
7, NATIONAL GALLERY, LONDON, U.K., 285%, 2727, 709, 1197, 6011
8, NATURAL HISTORY MUSEUM, LONDON, U.K., 198%, 4655, 1564, 1197, 5424
9, AMERICAN MUSEUM OF NATURAL HISTORY, NEW YORK, U.S., 0, 0, 0, 0, 5000
10, STATE HERMITAGE, ST. PETERSBURG, RUSSIA, 71%, 2813, 1649, 969, 4957
11, SHANGHAI SCIENCE AND TECHNOLOGY MUSEUM, SHANGHAI, CHINA, -79%, 494, 2368, 1351, 4824
12, REINA SOFÍA, MADRID, SPAIN, 84%, 3063, 1643, 1248, 4426
13, NATIONAL MUSEUM OF NATURAL HISTORY, WASHINGTON, U.S., 297%, 3900, 982, 573, 4200
14, NANJING MUSEUM, NANJING, CHINA, -21%, 1610, 2034, 1515, 4169
15, ZHEJIANG MUSEUM, HANGZHOU, CHINA, -6%, 874, 929, 1108, 4150
16, NATIONAL GALLERY OF ART, WASHINGTON, U.S., 91%, 3256, 1705, 730, 4074
17, VICTORIA & ALBERT MUSEUM, LONDON, U.K., 176%, 2370, 858, 872, 3921
18, CHINA SCIENCE TECHNOLOGY MUSEUM, BEIJING, CHINA, -39%, 1431, 2360, 690, 3891
19, NATIONAL PALACE MUSEUM, TAIPEI, TAIWAN, 149%, 1040, 417, 643, 3832
20, MUSÉE D'ORSAY, PARIS, FRANCE, 213%, 3270, 1044, 867, 3652"

top20_mu_ww = read.table(text = data_text_mu_ww, sep = ",", header = FALSE, quote = "", na.strings = "-")
colnames(top20_mu_ww) = c("id","Park_Name", "City", "Country", "%Change", "2022", "2021", "2020", "2019")
```


#### Clean the dataset for the top 20 museums Worldwide in 2019 to 2022
```{r, echo=TRUE}
top20_mu_ww_clean_df =
  top20_mu_ww |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Museum",
    Region = "Worldwide",
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```


#### Import museums data of North America 2019-2022
```{r, echo=TRUE}
data_text_mu_na = "1, THE METROPOLITAN MUSEUM OF ART, NEW YORK, NY, U.S., 64%, 3209, 1958, 1125, 6770
2, AMERICAN MUSEUM OF NATURAL HISTORY, NEW YORK, NY, U.S., 0, 0, 0, 0, 5000
3, NATIONAL MUSEUM OF NATURAL HISTORY, WASHINGTON, DC, U.S., 297%, 3900, 982, 573, 4200
4, NATIONAL GALLERY OF ART, WASHINGTON, DC, U.S., 91%, 3256, 1705, 730, 4074
5, NATIONAL AIR AND SPACE MUSEUM, WASHINGTON, DC, U.S., 2%, 419, 411, 267, 3200
6, NATIONAL MUSEUM OF AMERICAN HISTORY, WASHINGTON, DC, U.S., 216%, 1800, 570, 311, 2800
7, CALIFORNIA SCIENCE CENTER, LOS ANGELES, CA, U.S., 133%, 1694, 726, 271, 2200
8, NATIONAL MUSEUM OF AFRICAN AMERICAN HISTORY AND CULTURE, WASHINGTON, DC, U.S., 178%, 1093, 393, 368, 2000
9, SMITHSONIAN AMERICAN ART MUSEUM (WITH THE RENWICK GALLERY), WASHINGTON, DC, U.S., 162%, 1100, 420, 387, 2000
10, THE MUSEUM OF MODERN ART, NEW YORK, NY, U.S., 89%, 2190, 1161, 706, 1992
11, HOUSTON MUSEUM OF NATURAL SCIENCE, HOUSTON, TX, U.S., 40%, 1520, 1086, 0, 1974
12, NATIONAL PORTRAIT GALLERY, WASHINGTON, DC, U.S., 189%, 954, 330, 321, 1700
13, THE ART INSTITUTE OF CHICAGO, CHICAGO, IL, U.S., 38%, 1037, 700, 375, 1666
14, U.S. HOLOCAUST MEMORIAL MUSEUM, WASHINGTON, DC, U.S., 164%, 1184, 448, 0, 1600
15, FIELD MUSEUM OF NATURAL HISTORY, CHICAGO, IL, U.S., 68%, 1018, 605, 316, 1494
16, DENVER MUSEUM OF NATURE & SCIENCE, DENVER, CO, U.S., 20%, 1151, 959, 991, 1487
17, MUSEUM OF SCIENCE, BOSTON, MA, U.S., 93%, 1123, 583, 345, 1458
18, THE J. PAUL GETTY CENTER, LOS ANGELES, CA, U.S., 80%, 909, 504, 267, 1439
19, MUSEUM OF SCIENCE AND INDUSTRY, CHICAGO, IL, U.S., 45%, 1134, 780, 291, 1388
20, CALIFORNIA ACADEMY OF SCIENCES, SAN FRANCISCO, CA, U.S., 50%, 966, 646, 240, 1321"

top20_mu_na = read.table(text = data_text_mu_na, sep = ",", header = FALSE, quote = "", na.strings = "-")
colnames(top20_mu_na) = c("id","Park_Name", "City", "State", "Country", "%Change", "2022", "2021", "2020", "2019")
```


#### Clean the dataset for the top 20 museums North America in 2019 to 2022
```{r, echo=TRUE}
top20_mu_na_clean_df =
  top20_mu_na |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Museum",
    Region = "North America",
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```


#### Import museums data of Asia Pacific 2019-2022
```{r, echo=TRUE}
data_text_mu_apac = "1, NATIONAL MUSEUM OF CHINA, BEIJING, CHINA, -31%, 1631, 2378, 1600, 7390
2, SHANGHAI SCIENCE & TECHNOLOGY MUSEUM, SHANGHAI, CHINA, -79%, 494, 2368, 1351, 4824
3, NANJING MUSEUM, NANJING, CHINA, -21%, 1610, 2034, 1515, 4169
4, ZHEJIANG MUSEUM, HANGZHOU, CHINA, -6%, 874, 929, 1108, 4150
5, CHINA SCIENCE TECHNOLOGY MUSEUM, BEIJING, CHINA, -39%, 1431, 2360, 690, 3891
6, NATIONAL PALACE MUSEUM, TAIPEI, TAIWAN, 149%, 1040, 417, 643, 3832
7, XIAN MUSEUM, XIAN, CHINA, 27%, 401, 315, 348, 3420
8, NATIONAL MUSEUM OF KOREA, SEOUL, SOUTH KOREA, 170%, 3411, 1263, 774, 3354
9, TIANJIN NATURAL HISTORY MUSEUM, TIANJIN, CHINA, -17%, 1321, 1600, 550, 3150
10, HUNAN MUSEUM, CHANGSHA, CHINA, -14%, 1221, 1421, 1130, 3140
11, NATIONAL TAIWAN SCIENCE EDUCATION CENTER, TAIPEI, TAIWAN, 62%, 1879, 1158, 1082, 3000
12, SHANXI HISTORY MUSEUM, XIAN, CHINA, -54%, 690, 1510, 1000, 2900
13, TOKYO METROPOLITAN ART MUSEUM, TOKYO, JAPAN, 44%, 1510, 1049, 510, 2874
14, NATIONAL MUSEUM OF NATURAL SCIENCE, TAICHUNG, TAIWAN, 47%, 2300, 1566, 2263, 2650
15, SHANGHAI NATURAL HISTORY MUSEUM, SHANGHAI, CHINA, -75%, 532, 2166, 1040, 2549
16, NATIONAL MUSEUM OF NATURE AND SCIENCE, TOKYO, JAPAN, 69%, 1350, 800, 580, 2460
17, ZHEJIANG MUSEUM OF NATURAL HISTORY, HANGZHOU, CHINA, 0%, 738, 740, 600, 2443
18, NATIONAL GALLERY OF VICTORIA, MELBOURNE, AUSTRALIA, 97%, 1580, 802, 637, 2433
19, TOKYO NATIONAL MUSEUM, TOKYO, JAPAN, 127%, 1372, 605, 725, 2431
20, HUBEI PROVINCIAL MUSEUM, WUHAN, CHINA, 37%, 1000, 730, 446, 2400"

top20_mu_apac = read.table(text = data_text_mu_apac, sep = ",", header = FALSE, quote = "", na.strings = "-")
colnames(top20_mu_apac) = c("id","Park_Name", "City", "Country", "%Change", "2022", "2021", "2020", "2019")
```


#### Clean the dataset for the top 20 museums Asia Pacific in 2019 to 2022
```{r, echo=TRUE}
top20_mu_apac_clean_df =
  top20_mu_apac |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Museum",
    Region = "Asia Pacific",
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```


#### Import museums data of Europe Middle East Africa 2019-2022
```{r, echo=TRUE}
data_text_mu_emea = "1, LOUVRE, PARIS, FRANCE, 173%, 7726, 2825, 2700, 9600
2, VATICAN MUSEUMS, VATICAN, VATICAN CITY, 215%, 5081, 1613, 1300, 6883
3, BRITISH MUSEUM, LONDON, U.K., 209%, 4097, 1327, 1275, 6208
4, TATE MODERN, LONDON, U.K., 236%, 3883, 1156, 1433, 6098
5, NATIONAL GALLERY, LONDON, U.K., 285%, 2727, 709, 1197, 6011
6, NATURAL HISTORY MUSEUM, LONDON, U.K., 198%, 4655, 1564, 1197, 5424
7, STATE HERMITAGE, ST PETERSBURG, RUSSIA, 71%, 2813, 1649, 969, 4957
8, REINA SOFÍA, MADRID, SPAIN, 86%, 3063, 1643, 1248, 4426
9, VICTORIA & ALBERT MUSEUM, LONDON, U.K., 176%, 2370, 858, 872, 3921
10, MUSÉE D'ORSAY, PARIS, FRANCE, 213%, 3270, 1044, 867, 3652
11, CENTRE POMPIDOU, PARIS, FRANCE, 101%, 3010, 1501, 913, 3270
12, SCIENCE MUSEUM (SOUTH KENSINGTON), LONDON, U.K., 262%, 2335, 646, 862, 3254
13, MUSEO NACIONAL DEL PRADO, MADRID, SPAIN, 118%, 2457, 1127, 852, 3203
14, STATE TRETYAKOV GALLERY, MOSCOW, RUSSIA, 21%, 1910, 1581, 894, 2836
15, RIJKSMUSEUM, AMSTERDAM, NETHERLANDS, 183%, 1766, 625, 675, 2700
16, CITE DES SCIENCES ET DE L'INDUSTRIE, PARIS, FRANCE, 114%, 2101, 984, 793, 2370
17, GALLERIA DEGLI UFFIZI, FLORENCE, ITALY, 129%, 2223, 970, 659, 2362
18, AUSCHWITZ-BIRKENAU MUSEUM, OŚWIĘCIM, POLAND, 110%, 1184, 563, 502, 2300
19, NATIONAL MUSEUM OF SCOTLAND, EDINBURGH, U.K., 199%, 1974, 661, 444, 2210
20, VAN GOGH MUSEUM, AMSTERDAM, NETHERLANDS, 267%, 1346, 366, 517, 2100"

top20_mu_emea = read.table(text = data_text_mu_emea, sep = ",", header = FALSE, quote = "", na.strings = "-")
colnames(top20_mu_emea) = c("id","Park_Name", "City", "Country", "%Change", "2022", "2021", "2020", "2019")
```


#### Clean the dataset for the top 20 museums Europe Middle East Africa in 2019 to 2022
```{r, echo=TRUE}
top20_mu_emea_clean_df =
  top20_mu_emea |>
  pivot_longer(
    ("2019"):("2022"),
    names_to = "Year",
    values_to = "Attendance"
  ) |>
  mutate(
    Type = "Museum",
    Region = "Europe Middle East Africa",
    Attendance = Attendance * 10e3
  ) |>
  select(
    Park_Name, City, Country, Year, Attendance, Type, Region
  )
```


### Combine all dataset
```{r, echo=TRUE}
combine_df = 
  bind_rows(top25_tp_ww_clean_df, top20_tp_na_clean_df, top10_tp_la_clean_df, top20_tp_apac_clean_df, top20_tp_emea_clean_df, top20_wp_na_clean_df, top10_wp_la_clean_df, top20_wp_apac_clean_df, top10_wp_emea_clean_df, top20_mu_ww_clean_df, top20_mu_na_clean_df, top20_mu_apac_clean_df, top20_mu_emea_clean_df)
```


### Export the final dataset
```{r, echo=TRUE}
write.csv(combine_df, "ultimate data.csv", row.names = FALSE)
```

## Exploratory analysis

```{r, include = FALSE}
library(tidyverse)
library(plotly)
library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

After tidying our data set, an exploratory analysis is conducted to look for possible predictors for the `Attendance` outcome. 

```{r, include=FALSE}
theme_park = read_csv("ultimate data.csv") |> 
    mutate(
    Year = as.factor(Year),
    Type = as.factor(Type)
  ) |> 
    filter(
      Region != c("Worldwide")
    ) 
```

### Descriptive Analysis for All Data

#### Total Attendance by Type from 2019 to 2022

A brief summary of attendance based on the `Type` variable is provided below:

```{r region attend summary,message = FALSE, echo = FALSE}
theme_park  |> 
    group_by(Year, Type) |> 
    mutate(
      Attendance = Attendance / 100000
    ) |> 
    summarise(sum = sum(Attendance)) |> 
    arrange(Type) |> 
    pivot_wider(
      names_from = Type,
      values_from = sum
    ) |> 
    knitr::kable(digits = 3, caption = c("Summary of Attendance for Three Types of Facilities From 2019 to 2022"))
```

From this table, some observed patterns are:

* A common decreasing trend of attendance from 2019 and 2020
* A common increasing trend of attendance from 2020 and 2022

The distribution of data by year is further visualized into the box plots below:

```{r}
theme_park |> 
  group_by(Year) |> 
  plot_ly(y = ~Attendance, color = ~Year, type = "box", colors = "viridis") |> 
  layout(annotations = 
 list(x = 1, y =1, text = "Plot 1: Distribution of Attendance by Year", 
      showarrow = F, xref='paper', yref='paper', 
      xanchor='right', yanchor='auto', xshift=0, yshift=0,
      font=list(size=15))
 )
```

* Overall high attendance in 2019 and low attendance in 2020
* Common outliers in the distribution of attendance for each year

#### Total Attendance by Region from 2019 to 2022

Next, we specifically look at the trend of `Attendance` from 2019 to 2022 based on the `Region` variable. 

```{r, message = FALSE, echo = FALSE}
 theme_park|> 
  group_by(Region, Year)  |> 
  summarize(attend_sum = mean(Attendance)) |> 
  plot_ly(x = ~Year, y = ~attend_sum, color = ~Region,
          type = "scatter", mode = 'point', colors = "viridis") |> 
    layout(annotations = 
 list(x = 1, y = 1, text = "Plot 2: Change in Attendance for Each Region", 
      showarrow = F, xref='paper', yref='paper', 
      xanchor='right', yanchor='auto', xshift=0, yshift=0,
      font=list(size=15))
 )
```

* Similar attendance fluctuation for most regions
* Only EMEA has an attendance drop from 2021 to 2022

### Descriptive Analysis For All Data （Worldwide Top 25)

```{r, message = FALSE, echo = FALSE}
theme_full = 
  read_csv("ultimate data.csv")
```

#### Top 25 theme park from worldwide's sum attendance by different country.

Looking at the plot 1, we found that there was too much data for outliers, so we did another data analysis based on Top 25 theme park, which is labled as Worldwide on the region variable.
```{r}
#| echo: false
#| message: false
theme_full|> 
  filter(Region == "Worldwide") |> 
  group_by(Year) |> 
  plot_ly(x = ~ Year, y = ~ Attendance, color = ~ Country, type = "bar", colors = "viridis", alpha = 0.5) |> 
  layout(annotations = 
 list(x = 1, y =1.04, text = "Plot 3: Distribution of Attendance by different Country in different year", 
      showarrow = F, xref='paper', yref='paper', 
      xanchor='right', yanchor='auto', xshift=0, yshift=0,
      font=list(size=15))
 )
```

We use "Worldwide" to rank the top 25 theme parks by total number of visitors in different countries in different year. According to the icon, we can get:

* The total number of tourists from U.S. is the largest from 2019 to 2022, followed by China and Japan in third place.

* Russia, Germany and Spain have the lowest total tourist arrivals in 2019-2022.

* Based on comparisons between different years, we found that the number of tourists dropped sharply from 2019 to 2020. The number of tourists from 2020 to 2022 is slowly returning to normal, but it is not as high as before the epidemic.

### ANOVA test

#### Based on Type of Facilities

The first ANOVA test focuses on the `Type` variable in our data set. The null hypothesis and alternative hypothesis are listed as follow:

$$H_0: \mu_{\text{Amusement/Theme Park}} = \mu_{\text{Water Park}} = \mu_{\text{Museum}} ~~ \text{vs} ~~ H_1: \text{at least two means are not equal}$$
```{r}

anova_1 = aov(Attendance ~ Type, data = theme_park)

summary(anova_1)

```

With a p-value of less than 2e-16, we would reject the null hypothesis. We have evidence that at least
two of the means are not equal. Meaning the mean attendance among type groups is different
for at least two groups in the `Type` variable.

#### Based on thr Type of Year
The Second ANOVA test focuses on the `Year` variable in our data set. The null hypothesis and alternative hypothesis are listed as follow:

$$H_0: \mu_{\text{2019}} = \mu_{\text{2020}} = \mu_{\text{2021}}= \mu_{\text{2022}} ~~ \text{vs} ~~ H_1: \text{at least two means are not equal}$$
```{r}

dat = 
  theme_full |> 
  filter(
    Region != c("Worldwide")
    ) |> 
  mutate(
    Year = as.factor(Year)
  ) 

anova2 = aov(Attendance ~ Year , data = dat) 

summary(anova2)

```
With a p-value of less than 2e-16, we would reject the null hypothesis. We have evidence that at least two of the means are not equal. Meaning the mean attendance among year groups is different for at least two groups.

## Regression Analysis

```{r, include=FALSE}
library(tidyverse)
```

```{r, include=FALSE}
data_df =
  read_csv("ultimate data.csv") |>
  janitor::clean_names() |>
  select(attendance, everything()) |>
  filter(!region == "Worldwide")
```

### Linear Regression
Linear regression is used to model the relationship between attendance and five independent variables (city, country, year, type, and region). It assumes that this relationship is linear, meaning changes in the predictors are associated with a constant change in the response. We first fit linear regression of each predictors to check if they are associated with attendance.

#### Relationship between city and attendance
```{r, include=FALSE}
city_fit = lm(attendance ~ city, data = data_df) 
summary(city_fit)
```

```{r, message = FALSE, echo = FALSE}
data_df |>
  ggplot(aes(city, attendance)) + geom_point(color='pink') + 
  theme_bw() +
  geom_smooth(method='lm', se=FALSE, color='orange') +
  labs(x="City", y="Attendance ")
```

a few cities are associated with attendance such as ANAHEIM, Beijing, UNIVERSAL CITY, TOKYO, VATICAN. In general, there is no apparent linear relationship between city and attendance.

#### Relationship between country and attendance
```{r, include=FALSE}
country_fit = lm(attendance ~ country, data = data_df) 
summary(country_fit)
```

```{r, message = FALSE, echo = FALSE}
data_df |>
  ggplot(aes(country, attendance)) + geom_point(color='pink') + 
  theme_bw() +
  geom_smooth(method='lm', se=FALSE, color='orange') +
  labs(x="Country", y="Attendance ")
```

only one country Japan is significantly associated with attendance. In general, there is no apparent linear relationship between country and attendance. 

#### Relationship between year and attendance
```{r, message = FALSE, echo = FALSE}
year_fit = lm(attendance ~ year, data = data_df) 
summary(year_fit)
```

according to the summary of linear regression, $\widehat{Y}=\widehat{\beta_0}+\widehat{\beta_1}*X$.
Linear Model: $\widehat{attendance}=6104922321-3011254*year$.

```{r, message = FALSE, echo = FALSE}
data_df |>
  ggplot(aes(year, attendance)) + geom_point(color='pink') + 
  theme_bw() +
  geom_smooth(method='lm', se=FALSE, color='orange') +
  labs(x="Year", y="Attendance ")
```

Year is significantly associated with attendance, and there is a linear relationship between year and attendance. Attendance has experienced a decline from 2019 to 2022. 

#### Relationship between park type and attendance
```{r, include=FALSE}
type_fit = lm(attendance ~ type, data = data_df) 
summary(type_fit)
```

```{r, message = FALSE, echo = FALSE}
data_df |>
  ggplot(aes(type, attendance)) + geom_point(color='pink') + 
  theme_bw() +
  geom_smooth(method='lm', se=FALSE, color='orange') +
  labs(x="Type", y="Attendance ")
```

Park type is significantly associated with attendance. Amusement/Theme Park has the highest attendance and water park has the lowest attendance. 

#### Relationship between region and attendance
```{r, include=FALSE}
region_fit = lm(attendance ~ region, data = data_df) 
summary(region_fit)
```

```{r, message = FALSE, echo = FALSE}
data_df |>
  ggplot(aes(region, attendance)) + geom_point(color='pink') + 
  theme_bw() +
  geom_smooth(method='lm', se=FALSE, color='orange') +
  labs(x="region", y="Attendance ")
```

Latin America is significantly associated with attendance. Overall, there is no apparent linear relationship between region and attendance.  

#### Fit a Multiple linear regression
Multiple linear regression allows us to account for the influence of multiple independent variables on the attendance simultaneously. The outcomes of the linear regression analysis suggest that the variable "year" exhibits a linear relationship with attendance. Additionally, specific categories within the "type" and "region" variables are found to be significantly associated with attendance. Thus, variables "year", "type", and "region" are used to fit a multiple linear regression. 
```{r, echo = FALSE}
multi_fit = lm(attendance ~ year + type + region, data = data_df)
summary(multi_fit)
```

only North America within region is not significantly associated with attendance.

#### Stepwise regression
stepwise regression helps streamline the modeling process by automatically including or excluding variables based on statistical criteria. We use both backward and forward selection to find the best fit model. 
```{r, echo = FALSE}
step(multi_fit, direction='both')
```

The best fit model with lowest AIC is $\widehat{attendance}=\widehat{\beta_0}+\widehat{\beta_1}*year+\widehat{\beta_2}*type+\widehat{\beta_3}*region$.

### Logistic Regression

#### Background

In order to further study the relationship between different stages of the COVID-19 pandemic and the visitor flow of the theme park, based on the development process of the pandemic, the years from 2019 to 2022 are divided into two pandemic levels

* Outbreak period: 2019 to 2021

* Control period: 2022

Then add to our data as `pandemic_level`.

Furthermore, to classify the pandemic period by park attendance and other covariates, including Park Type and Region, we established a logistic regression model with `pandemic_level` as the dependent variable.

#### Data Preparation

```{r include=FALSE, message=FALSE}
library(tidyverse)
library(modelr)
library(viridis)
library(caret)
library(pROC)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 9, 
  fig.height = 6,
  out.width = "90%",
	fig.align = 'center'
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


```{r message=FALSE, warning=FALSE}
#import the data and tidy
set.seed(12138)

park_df =
  read_csv("ultimate data.csv") |>
  janitor::clean_names() |>
  drop_na()

logistic_df = 
  park_df |>
  mutate(
    type = as.factor(type),
    region = as.factor(region),
    country = as.factor(country),
    pandemic_level = case_when(
      year == 2019 ~ "outbreak",
      year == 2020 ~ "outbreak",
      year == 2021 ~ "outbreak",
      year == 2022 ~ "control"
    ),
    pandemic_level = as.factor(pandemic_level)
  ) 
```


```{r}
#attendance relationship
logistic_df |>
  ggplot(aes(park_name, log10(attendance), color = pandemic_level)) +
  geom_point(alpha = .7) +
  facet_grid(. ~ region ) +
  labs(x = "Parks", y = "Attendance (log10)",
       caption = "Fig.1 The relationship between park attendance and pandemic level across regions") +
  theme(axis.text.x = element_blank() )
```

```{r}
#split data into train and test subset
cv_results = 
  logistic_df |>
  filter(region != "Worldwide") |>
  mutate(
    pandemic_level = 
      case_match(
        pandemic_level,
        "outbreak" ~ 1,
        "control" ~ 0
      )
  )

cv_df = 
  crossv_mc(
  cv_results,1
)

cv_df =
  cv_df |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))
```

After transforming attendance by log function, the relationship between the park attendance an the pandemic levels was displayed in Fig.1. 

It can be noticed that different epidemic stages may have different visitor flow patterns, in other words, the traffic of theme parks from 2019 to 2022 can be classified according to the extent of the epidemic. This is particularly true in EMEA, the Middle East and Latin America.

Then the observations in region *"Worldwide"* were excluded as redundant data and `pandemic_level` is converted to a binomial distribution in order to fit logistic regression models and predict epidemic levels more easily.

#### Logistic model

```{r results='hide'}
#fit model and selection
log_mod =  
  cv_df |>
  select(train) |>
  unnest(cols = c(train)) |>
  glm(pandemic_level ~ attendance + type + region, family = "binomial", data = _)

step(log_mod, direction = "forward") 

best_fit = log_mod
```

```{r}
#model summary
coef(best_fit) |>
  knitr::kable(col.names = c("Variables" , "Coefficients"),
               caption = "Table.1 The coefficients of model")

best_fit |>
  broom::tidy() |>
  knitr::kable(caption = "Table.2 Summary of the chose model" )

best_fit |>
  broom::glance() |>
  knitr::kable(caption = "Table.3 The AIC of the model")
```

A full logistic regression model was fitted to our data with all Concomitant variables that we were interested in, including `attendance`, `type`, and `region` by `glm()` function with `binomial` as family.

Then forward step-wise model selection was performed by `step()` to selected the best fitted model with lowest AIC value. 

After step-wise selection, we chose the full model as the final logistic regression model while `attendance`, `typeMuseum`, `regionEMEA` and `regionLatin America` showed significant association with the `pandemic_level`. As showed in Table.1, the model statement is \newline

*pandemic_level = 2.2365446 + (-8.5e-09)attendance + (-0.8417669)Museum + (-0.5368024)Water Park + (-0.7424560)EMEA + (-0.0950834)Europe Middle East Africa + (-0.8266498)Latin America +(-0.4789397)North America*

#### Cross validation and ROC

```{r}
#cross validation 5 fold
set.seed(1)
folds <- createFolds(y=pull(cv_results, pandemic_level),k=5)

#five fold ROC
auc_value<-as.numeric()

for(i in 1:5){
  fold_test <- cv_results[folds[[i]],] 
  fold_train <- cv_results[-folds[[i]],] 
  fold_pre <- glm(pandemic_level ~ attendance + type + region, family = "binomial", data = fold_train )
  fold_predict <- predict(fold_pre,type='response',newdata=fold_test)
  auc_value<- append(auc_value,as.numeric(auc(as.numeric(pull(fold_test, pandemic_level)),fold_predict)))
}


#ROC for fold 1 as example
fold_test <- cv_results[folds[[1]],]
fold_train <- cv_results[-folds[[1]],]

fold_pre <- glm(pandemic_level ~ attendance + type + region, family = "binomial", data = fold_train ) 
fold_predict <- predict(fold_pre,type='response',newdata=fold_test)
roc1<-roc(pull(fold_test, pandemic_level),fold_predict)

plot(roc1, print.auc=T, auc.polygon=T, grid=c(0.1, 0.2),
     grid.col=c("green", "red"), max.auc.polygon=T,
     auc.polygon.col="skyblue", 
     print.thres=T)
mtext("Fig.3 The ROC curve for fold 1 of the model", side = 1, line = 4.1)
``` 

After 5 fold cross validation, the model performance was indicated by a mean AUC value, which equals to `r mean(auc_value)`, the ROC curve for fold 1 was displayed in Fig.3 as an example. 

#### Disscusion

According to the results, the prediction ability of the model was not excellent with a low AUC value, which may cause by the small sample size. In addition, in the binary classification of dependent variables, the sample size gap between the two categories was large, and the model can be improved by under-sampling or over-sampling methods. Moreover, the prediction ability of the model can be improved by adding weight parameters when establishing the model.

## Discussion

### Result Summary

Our in-depth analysis of theme park attendance involved the application of linear and logistic regression models, exploratory data analysis, and ANOVA tests. The linear regression delved into the connection between attendance and diverse predictors, including city, country, year, type, and region. This examination unearthed significant associations between attendance and specific cities, countries, years, park types, and regions. Subsequently, we implemented multiple linear regression, incorporating variables that we interested in, with stepwise regression fine-tuning the model. \newline

The logistic regression model introduced a new variable, `pandemic_level`, reflecting different stages of the COVID-19 pandemic. This logistic regression model successfully identified noteworthy associations between attendance, park type, region, and the pandemic level. Our analytical progress encompassed data preparation, cross-validation, ROC curve visualization, and detailed discussions on model performance. Descriptive analyses provided concise summaries of attendance trends categorized by type, region, and country, offering valuable insights into the global landscape of theme park attendance. \newline

Moreover, ANOVA tests were employed to rigorously evaluate the significance of mean differences in attendance across various types and years. This comprehensive analysis seamlessly integrated statistical techniques and visualization, resulting in a thorough exploration of the theme park dataset.

### Discussion

Furthermore, there are still many limitations in our study, such as small sample size and insufficient sample level, which may lead to unclear trends of park flow in different regions or types during visualization, as well as low sensitivity of regression models. \newline

In future studies, other factors that may be related to the flow of visitors to theme parks, such as park size, opening policies, and annual income of facilities, can be added, in order to further explore the overall impact of the epidemic on theme parks. At the same time, under-sampling or over-sampling methods can be applied to improve the balance of data, which may improve the prediction ability and accuracy of the regression models.