---
title: "Exploratory Data Analysis"
author: "Zixuan Zhang, Yilei Yang"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---
```{r, include = FALSE}
library(tidyverse)
library(plotly)
library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

After tidying our data set, an exploratory analysis is conducted to look for possible predictors for the `Attendance` outcome. 

```{r, include=FALSE}
theme_park = read_csv("ultimate data.csv") |> 
    mutate(
    Year = as.factor(Year),
    Type = as.factor(Type)
  ) |> 
    filter(
      Region != c("Worldwide")
    ) 
```

## Descriptive Analysis for All Data

### Total Attendance by Type from 2019 to 2022

A brief summary of attendance based on the `Type` variable is provided below:

```{r region attend summary,message = FALSE, echo = FALSE}
theme_park  |> 
    group_by(Year, Type) |> 
    mutate(
      Attendance = Attendance / 100000
    ) |> 
    summarise(sum = sum(Attendance)) |> 
    arrange(Type) |> 
    pivot_wider(
      names_from = Type,
      values_from = sum
    ) |> 
    knitr::kable(digits = 3, caption = c("Summary of Attendance for Three Types of Facilities From 2019 to 2022"))
```

From this table, some observed patterns are:

* A common decreasing trend of attendance from 2019 and 2020
* A common increasing trend of attendance from 2020 and 2022

The distribution of data by year is further visualized into the box plots below:

```{r}
theme_park |> 
  group_by(Year) |> 
  plot_ly(y = ~Attendance, color = ~Year, type = "box", colors = "viridis") |> 
  layout(annotations = 
 list(x = 1, y =1, text = "Plot 1: Distribution of Attendance by Year", 
      showarrow = F, xref='paper', yref='paper', 
      xanchor='right', yanchor='auto', xshift=0, yshift=0,
      font=list(size=15))
 )
```

* Overall high attendance in 2019 and low attendance in 2020
* Common outliers in the distribution of attendance for each year

### Total Attendance by Region from 2019 to 2022

Next, we specifically look at the trend of `Attendance` from 2019 to 2022 based on the `Region` variable. 

```{r, message = FALSE, echo = FALSE}
 theme_park|> 
  group_by(Region, Year)  |> 
  summarize(attend_sum = mean(Attendance)) |> 
  plot_ly(x = ~Year, y = ~attend_sum, color = ~Region,
          type = "scatter", mode = 'point', colors = "viridis") |> 
    layout(annotations = 
 list(x = 1, y = 1, text = "Plot 2: Change in Attendance for Each Region", 
      showarrow = F, xref='paper', yref='paper', 
      xanchor='right', yanchor='auto', xshift=0, yshift=0,
      font=list(size=15))
 )
```

* Similar attendance fluctuation for most regions
* Only EMEA has an attendance drop from 2021 to 2022

## Descriptive Analysis For All Data （Worldwide Top 25)

```{r, message = FALSE, echo = FALSE}
theme_full = 
  read_csv("ultimate data.csv")
```

### Top 25 theme park from worldwide's sum attendance by different country.

Looking at the plot 1, we found that there was too much data for outliers, so we did another data analysis based on Top 25 theme park, which is labled as Worldwide on the region variable.
```{r}
#| echo: false
#| message: false
theme_full|> 
  filter(Region == "Worldwide") |> 
  group_by(Year) |> 
  plot_ly(x = ~ Year, y = ~ Attendance, color = ~ Country, type = "bar", colors = "viridis", alpha = 0.5) |> 
  layout(annotations = 
 list(x = 1, y =1.04, text = "Plot 3: Distribution of Attendance by different Country in different year", 
      showarrow = F, xref='paper', yref='paper', 
      xanchor='right', yanchor='auto', xshift=0, yshift=0,
      font=list(size=15))
 )
```

We use "Worldwide" to rank the top 25 theme parks by total number of visitors in different countries in different year. According to the icon, we can get:

* The total number of tourists from U.S. is the largest from 2019 to 2022, followed by China and Japan in third place.

* Russia, Germany and Spain have the lowest total tourist arrivals in 2019-2022.

* Based on comparisons between different years, we found that the number of tourists dropped sharply from 2019 to 2020. The number of tourists from 2020 to 2022 is slowly returning to normal, but it is not as high as before the epidemic.



## ANOVA test

### Based on Type of Facilities

The first ANOVA test focuses on the `Type` variable in our data set. The null hypothesis and alternative hypothesis are listed as follow:

$$H_0: \mu_{\text{Amusement/Theme Park}} = \mu_{\text{Water Park}} = \mu_{\text{Museum}} ~~ \text{vs} ~~ H_1: \text{at least two means are not equal}$$
```{r}

anova_1 = aov(Attendance ~ Type, data = theme_park)

summary(anova_1)

```

With a p-value of less than 2e-16, we would reject the null hypothesis. We have evidence that at least
two of the means are not equal. Meaning the mean attendance among type groups is different
for at least two groups in the `Type` variable.

### Based on thr Type of Year
The Second ANOVA test focuses on the `Year` variable in our data set. The null hypothesis and alternative hypothesis are listed as follow:

$$H_0: \mu_{\text{2019}} = \mu_{\text{2020}} = \mu_{\text{2021}}= \mu_{\text{2022}} ~~ \text{vs} ~~ H_1: \text{at least two means are not equal}$$
```{r}

dat = 
  theme_full |> 
  filter(
    Region != c("Worldwide")
    ) |> 
  mutate(
    Year = as.factor(Year)
  ) 

anova2 = aov(Attendance ~ Year , data = dat) 

summary(anova2)

```
With a p-value of less than 2e-16, we would reject the null hypothesis. We have evidence that at least two of the means are not equal. Meaning the mean attendance among year groups is different for at least two groups.
